### 2019-01-06 Paper/Manuscript.Rmd
first example doc uploaded

---
title: An R markdown template for Elsevier journals with examples for beginners 
author:
  - name: Paul Raftery 
    email: p.raftery@berkeley.edu
    affiliation: CBE
    footnote: Corresponding Author
  - name: Next contributor 
    affiliation: Organisation2
address:
  - code: CBE
    address: Center for the Built Environment, UC Berkeley, 390 Wurster Hall, Berkeley, CA, 94720, USA
  - code: Organisation2
    address: Another organization, and their address

abstract: Enter the text for your abstract here
journal: "Building and Environment"
date: "`r Sys.Date()`"
bibliography: Bibliography.bib
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
csl: elsevier-with-titles.csl
geometry: "left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm"
header-includes:
- \usepackage{setspace}
- \doublespacing
---

``` {r setup, echo=FALSE, cache=FALSE, results="hide", message=FALSE}
## code chunk for libraries and global settings (e.g. applies to the whole markdown file)
require(pander)
require(tidyverse)
require(ggExtra)
require(gridExtra)
require(here)
require(readxl)
require(broom)
require(knitr)
require(stargazer)
require(RColorBrewer)
require(ggrepel)
require(ggpmisc)

## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 1, digits = 2)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, fig.pos = 'h')

# set the gplot theme
theme_set(theme_light())
# update theme to lighten grey background, use black text, and 
# rotate text appropriately for facets as default
theme_update(strip.background = element_rect(fill="#F5F5F5", color="grey"),
             # remove background entirely with strip.background = element_blank(),
             strip.text = element_text(color = 'black'),
             strip.text.y = element_text(angle = 90))

# 2 print and color blind safe color options from UC Berkeley's official color palette.
cal_palette2 <- c("#3B7EA1", "#C4820E")
# 4 color version, with slight mod to original dark yellow for better grayscale performance 
cal_palette4 <- c("#003262", "#b0750c", "#3b7ea1", "#fdb515")
```

``` {r readSourceData, results="hide"}
#read your source dataset(s) from subfolder
# df <- read_csv(here("Paper","SupplementaryMaterial","RawData","all_test_data_and_metadata.csv"), guess_max=100000)
```

``` {r preprocessData, echo=FALSE, results="hide"}
# here is where all the preprocessing happens
#
# ensure you comment any sections or functions in sufficient detail that
# their purpose can be readily understood without reading through the detail 
# of the code itself
```

Keywords\:  
Key 1\; Key 2\; Key 3\; Key 4\; Key 5\; Key 6 (max)

\pagebreak

**Highlights\:**

-   Add your 3 -5 highlights
-   As bullets here
-   Making sure each is less than 85 characters in length

**Graphical Abstract**  
```{r graphabs, message=FALSE}
#include a graphical abstract, without a caption
include_graphics(here("Paper","SupplementaryMaterial","Figures","Latex_logo.png"))

# note the use of the here package (https://github.com/r-lib/here). This package allows you to define any filepath relative to an (empty) .here file, which means no more messing around changing filepaths!

# If there are spaces in your filepath above, the filepath prints just before the picture. While there are other ways to avoid this, the easy solution is not to use spaces in your folders or files.
```

\pagebreak

# Introduction
The purpose of this (very much work-in-progress) document is to provide a complete R markdown template for an Elsevier journal submission (based on the rticles repository @LaTeXJournalArticle2019), along with useful examples and packages to improve usability for folks who are just starting out with this workflow. The eventual intent is to capture minimal examples of the common things that authors need to do when writing papers in R markdown; provide examples of useful packages, workflows, and tools; and provide solutions to common issues that folks encounter.

# Examples
# Heading level 1
## Heading level 2
### Heading level 3
Here's how to **bold** or *italic* a piece of text.

This is how you do a bullet point list:

* First bullet
* Second bullet
    + A sub-bullet
    + Another sub-bullet

Or an ordered option:

1. Item 1
2. Item 2
    + Item 2a
    + Item 2b

## Citations
Citing other literature is remarkably easy, just like this @coakleyReviewMethodsMatch2014. This citation key references the tag associated with an entry in Bibliography.bib (a BibTex file). I've found it easiest to use Zotero to manage my library of references and to generate the BibTex file, though any software that creates a valid BibTex file should work fine. Zotero allows you to create a 'Collection' (or folder) that gathers together all of the references used for a particular document. When combined with with the Better BibTex plugin, that collection can be exported to a BibTex file that is continually updated as you add or edit references in that Collection. Better BibTex also puts the citation key - the text after the '\@' sybmbol in the .Rmd file - on the upper right of each entry, which is convenient for adding citations.

There's not much else involved in citing, as the references list gets built and formatted automatically based on the selected style. The only other issue I've had to look around to solve was figuring out how to combine multiple citations, which is easy when you know how. [@coakleyReviewMethodsMatch2014; @zhaiHumanComfortPerceived2015a]

## Cross-referencing
This is how you refer to a figure in your text: Figure \@ref(fig:correlation). Simply reference the title of the code chunk, and ensure that the code chunk includes a figure caption.

```{r correlation, fig.width = 1, fig.cap = "Correlation. Source: XKCD, xkcd.com/552"}
include_graphics(here("Paper","SupplementaryMaterial","Figures","Correlation.png"), dpi= 500)
```

## Calculations in text
The holy grail of markdown - doing all of your calculations in the same file so you never need to worry about updating them after someone[^footnoteApology] asks you to make changes.... again! It's as easy as pi: `r pi`. Incidentally,you can selectively override the 'global' options set at the beginning, to say for example, show more decimals: `r format(pi, digits=5)`.

[^footnoteApology]: Often I'm the someone, sorry CBE grad students. Also, look, it's an example of a footnote!

``` {r calculationOnlyChunk, results="hide"}
# do some calcs here, but don't output anything to the reader, 
# simply save the result for later use or reference in text
xyz = 18*2
```

This is an example of outputing the result of a calculation that you perform within a code chunk in the document somewhere prior to the location where you first refer to it: `r xyz`.

## Other packages
Here are lots of packages that are useful for markdown docs. I encourage you to search for these whenever you encounter a new thing you need to do and to propose an addition to this repository accordingly. For example, the Stargazer package describes linear models in a nice table.

``` {r linmods, results='asis', echo=FALSE, warning=FALSE}
lm1 = lm(hwy ~ displ + cyl + year, mpg)
lm2 = lm(cty ~ displ + cyl + year, mpg)
# uncomment next line for printing in console (i.e. human readable in RStudio)
#stargazer(lm1, lm2, type="text")
stargazer(lm1, lm2, dep.var.labels=c("Highway","City"), covariate.labels=c("Displacement","Num. cylinders", "Year"), title = "Car miles per gallon", column.sep.width = "1pt", header=FALSE)
```

## Equations and math
Here's a basic example inline $example_{subscript} = \frac{D}{R}$, or you display it on a whole line if needed. Google latex math cheat sheets for more information.

$$\sum_{i=1}^{n}{x_i^2}$$

# Writing style
This is a little off topic for an Rmd example but a convenient place to remind our grad students about writing style. In almost all cases, active voice is better than passive voice. Several psychological studies show that the active voice is more easily understood by readers, and that information is more accurately reported by authors when writing in active voice. For example, research @klenbortMarkednessPerspectiveInterpretation1974 has shown that the “active [voice] offers a neutral structure for conveying information”. Authorship guides for highly regarded journals often indicate a preference for the active voice instead of passive:

+  Nature: "Nature journals like authors to write in the active voice ('we performed the experiment...') as experience has shown that readers find concepts and results to be conveyed more clearly if written directly.”@NatureHowWrite
+  Science: "Use active voice when suitable, particularly when necessary for correct syntax (e.g., 'To address this possibility, we constructed a lZap library ...,' not 'To address this possibility, a lZap library was constructed...')." @rubenHowWriteScientist2012

And, on top of all that, you also end up with less text if you write in active voice, saving space for useful information and making it easier for your readers to understand.

## A fun way to spot passive voice:
If you can add the words ‘by zombies’ @ScaryeasyWayHelp2014 to the end of the sentence and the sentence still makes logical sense, then the sentence is in passive voice. You can also switch on the grammar settings in Microsoft Word’s spelling and grammar checker and it will show up that way.

How to fix it?

Change:

“These measurements are not quantitatively reported in the paper” (... by zombies)

To

“The paper does not quantitatively report these measurements”

Or even better, it's really the authors doing the reporting as the paper is an inanimate object...

“We do not quantitatively report these measurements”.

Change:

“Six different table and partition configurations were tested” (... by zombies)

To

“We tested six different table and partition configurations.”

# Methods

# Results

# Discussion

# Conclusion

# Acknowledgements
Don't forget to acknowledge the funder(s) with associated grant numbers if required. The same goes for folks who significantly assisted you with this paper but that are not authors.

# Declaration of interest
Descibe any relevant interests of the authors, particularly if there is a link to the research that is relatively uncommon and could be perceived as a conflict of interest.

# References {#references .unnumbered}

### 2019-01-06 Paper/Manuscript.tex
first example doc uploaded

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Building and Environment} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={An R markdown template for Elsevier journals with examples for beginners},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)
% Pandoc header
\usepackage{setspace}
\doublespacing

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\begin{frontmatter}

  \title{An R markdown template for Elsevier journals with examples for beginners}
    \author[CBE]{Paul Raftery\corref{c1}}
   \ead{p.raftery@berkeley.edu} 
   \cortext[c1]{Corresponding Author}
    \author[Organisation2]{Next contributor}

      \address[CBE]{Center for the Built Environment, UC Berkeley, 390 Wurster Hall,
Berkeley, CA, 94720, USA}
    \address[Organisation2]{Another organization, and their address}

  \begin{abstract}
  Enter the text for your abstract here
  \end{abstract}

 \end{frontmatter}

Keywords:\\
Key 1; Key 2; Key 3; Key 4; Key 5; Key 6 (max)

\pagebreak

\textbf{Highlights:}

\begin{itemize}
\tightlist
\item
  Add your 3 -5 highlights
\item
  As bullets here
\item
  Making sure each is less than 85 characters in length
\end{itemize}

\textbf{Graphical Abstract}\\
\includegraphics{C:/Users/praft/Documents/GitHub/rmd-example/Paper/SupplementaryMaterial/Figures/Latex_logo.png}

\pagebreak

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The purpose of this (very much work-in-progress) document is to provide
a complete R markdown template for an Elsevier journal submission (based
on the rticles repository {[}1{]}), along with useful examples and
packages to improve usability for folks who are just starting out with
this workflow. The eventual intent is to capture minimal examples of the
common things that authors need to do when writing papers in R markdown;
provide examples of useful packages, workflows, and tools; and provide
solutions to common issues that folks encounter.

\hypertarget{examples}{%
\section{Examples}\label{examples}}

\hypertarget{heading-level-1}{%
\section{Heading level 1}\label{heading-level-1}}

\hypertarget{heading-level-2}{%
\subsection{Heading level 2}\label{heading-level-2}}

\hypertarget{heading-level-3}{%
\subsubsection{Heading level 3}\label{heading-level-3}}

Here's how to \textbf{bold} or \emph{italic} a piece of text.

This is how you do a bullet point list:

\begin{itemize}
\tightlist
\item
  First bullet
\item
  Second bullet

  \begin{itemize}
  \tightlist
  \item
    A sub-bullet
  \item
    Another sub-bullet
  \end{itemize}
\end{itemize}

Or an ordered option:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Item 1
\item
  Item 2

  \begin{itemize}
  \tightlist
  \item
    Item 2a
  \item
    Item 2b
  \end{itemize}
\end{enumerate}

\hypertarget{citations}{%
\subsection{Citations}\label{citations}}

Citing other literature is remarkably easy, just like this {[}2{]}. This
citation key references the tag associated with an entry in
Bibliography.bib (a BibTex file). I've found it easiest to use Zotero to
manage my library of references and to generate the BibTex file, though
any software that creates a valid BibTex file should work fine. Zotero
allows you to create a `Collection' (or folder) that gathers together
all of the references used for a particular document. When combined with
with the Better BibTex plugin, that collection can be exported to a
BibTex file that is continually updated as you add or edit references in
that Collection. Better BibTex also puts the citation key - the text
after the `@' sybmbol in the .Rmd file - on the upper right of each
entry, which is convenient for adding citations.

There's not much else involved in citing, as the references list gets
built and formatted automatically based on the selected style. The only
other issue I've had to look around to solve was figuring out how to
combine multiple citations, which is easy when you know how. {[}2,3{]}

\hypertarget{cross-referencing}{%
\subsection{Cross-referencing}\label{cross-referencing}}

This is how you refer to a figure in your text: Figure
\ref{fig:correlation}. Simply reference the title of the code chunk, and
ensure that the code chunk includes a figure caption.

\begin{figure}
\centering
\includegraphics{C:/Users/praft/Documents/GitHub/rmd-example/Paper/SupplementaryMaterial/Figures/Correlation.png}
\caption{\label{fig:correlation}Correlation. Source: XKCD, xkcd.com/552}
\end{figure}

\hypertarget{calculations-in-text}{%
\subsection{Calculations in text}\label{calculations-in-text}}

The holy grail of markdown - doing all of your calculations in the same
file so you never need to worry about updating them after
someone\footnote{Often I'm the someone, sorry CBE grad students. Also,
  look, it's an example of a footnote!} asks you to make
changes\ldots{}. again! It's as easy as pi: 3.14. Incidentally,you can
selectively override the `global' options set at the beginning, to say
for example, show more decimals: 3.1416.

This is an example of outputing the result of a calculation that you
perform within a code chunk in the document somewhere prior to the
location where you first refer to it: 36.

\hypertarget{other-packages}{%
\subsection{Other packages}\label{other-packages}}

Here are lots of packages that are useful for markdown docs. I encourage
you to search for these whenever you encounter a new thing you need to
do and to propose an addition to this repository accordingly. For
example, the Stargazer package describes linear models in a nice table.

\begin{table}[!htbp] \centering 
  \caption{Car miles per gallon} 
  \label{} 
\begin{tabular}{@{\extracolsep{1pt}}lcc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
\cline{2-3} 
\\[-1.8ex] & Highway & City \\ 
\\[-1.8ex] & (1) & (2)\\ 
\hline \\[-1.8ex] 
 Displacement & $-$2.100$^{***}$ & $-$1.300$^{***}$ \\ 
  & (0.520) & (0.340) \\ 
  & & \\ 
 Num. cylinders & $-$1.300$^{***}$ & $-$1.200$^{***}$ \\ 
  & (0.410) & (0.270) \\ 
  & & \\ 
 Year & 0.150$^{***}$ & 0.071$^{**}$ \\ 
  & (0.055) & (0.036) \\ 
  & & \\ 
 Constant & $-$259.000$^{**}$ & $-$114.000 \\ 
  & (109.000) & (72.000) \\ 
  & & \\ 
\hline \\[-1.8ex] 
Observations & 234 & 234 \\ 
R$^{2}$ & 0.620 & 0.670 \\ 
Adjusted R$^{2}$ & 0.610 & 0.670 \\ 
Residual Std. Error (df = 230) & 3.700 & 2.500 \\ 
F Statistic (df = 3; 230) & 124.000$^{***}$ & 158.000$^{***}$ \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\hypertarget{equations-and-math}{%
\subsection{Equations and math}\label{equations-and-math}}

Here's a basic example inline \(example_{subscript} = \frac{D}{R}\), or
you display it on a whole line if needed. Google latex math cheat sheets
for more information.

\[\sum_{i=1}^{n}{x_i^2}\]

\hypertarget{writing-style}{%
\section{Writing style}\label{writing-style}}

This is a little off topic for an Rmd example but a convenient place to
remind our grad students about writing style. In almost all cases,
active voice is better than passive voice. Several psychological studies
show that the active voice is more easily understood by readers, and
that information is more accurately reported by authors when writing in
active voice. For example, research {[}4{]} has shown that the ``active
{[}voice{]} offers a neutral structure for conveying information''.
Authorship guides for highly regarded journals often indicate a
preference for the active voice instead of passive:

\begin{itemize}
\tightlist
\item
  Nature: ``Nature journals like authors to write in the active voice
  (`we performed the experiment\ldots{}') as experience has shown that
  readers find concepts and results to be conveyed more clearly if
  written directly.''{[}5{]}
\item
  Science: ``Use active voice when suitable, particularly when necessary
  for correct syntax (e.g., `To address this possibility, we constructed
  a lZap library \ldots{},' not `To address this possibility, a lZap
  library was constructed\ldots{}').'' {[}6{]}
\end{itemize}

And, on top of all that, you also end up with less text if you write in
active voice, saving space for useful information and making it easier
for your readers to understand.

\hypertarget{a-fun-way-to-spot-passive-voice}{%
\subsection{A fun way to spot passive
voice:}\label{a-fun-way-to-spot-passive-voice}}

If you can add the words `by zombies' {[}7{]} to the end of the sentence
and the sentence still makes logical sense, then the sentence is in
passive voice. You can also switch on the grammar settings in Microsoft
Word's spelling and grammar checker and it will show up that way.

How to fix it?

Change:

``These measurements are not quantitatively reported in the paper''
(\ldots{} by zombies)

To

``The paper does not quantitatively report these measurements''

Or even better, it's really the authors doing the reporting as the paper
is an inanimate object\ldots{}

``We do not quantitatively report these measurements''.

Change:

``Six different table and partition configurations were tested''
(\ldots{} by zombies)

To

``We tested six different table and partition configurations.''

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

Don't forget to acknowledge the funder(s) with associated grant numbers
if required. The same goes for folks who significantly assisted you with
this paper but that are not authors.

\hypertarget{declaration-of-interest}{%
\section{Declaration of interest}\label{declaration-of-interest}}

Descibe any relevant interests of the authors, particularly if there is
a link to the research that is relatively uncommon and could be
perceived as a conflict of interest.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-LaTeXJournalArticle2019}{}%
{[}1{]} LaTeX Journal Article Templates for R Markdown. Contribute to
rstudio/rticles development by creating an account on GitHub, (2019).

\leavevmode\hypertarget{ref-coakleyReviewMethodsMatch2014}{}%
{[}2{]} D. Coakley, P. Raftery, M. Keane, A review of methods to match
building energy simulation models to measured data, Renewable and
Sustainable Energy Reviews. 37 (2014) 123--141.
doi:\href{https://doi.org/10.1016/j.rser.2014.05.007}{10.1016/j.rser.2014.05.007}.

\leavevmode\hypertarget{ref-zhaiHumanComfortPerceived2015a}{}%
{[}3{]} Y. Zhai, Y. Zhang, H. Zhang, W. Pasut, E. Arens, Q. Meng, Human
comfort and perceived air quality in warm and humid environments with
ceiling fans, Building and Environment. 90 (2015) 178--185.
doi:\href{https://doi.org/10.1016/j.buildenv.2015.04.003}{10.1016/j.buildenv.2015.04.003}.

\leavevmode\hypertarget{ref-klenbortMarkednessPerspectiveInterpretation1974}{}%
{[}4{]} I. Klenbort, M. Anisfeld, Markedness and perspective in the
interpretation of the active and passive voice, Quarterly Journal of
Experimental Psychology. 26 (1974) 189--195.
doi:\href{https://doi.org/10.1080/14640747408400404}{10.1080/14640747408400404}.

\leavevmode\hypertarget{ref-NatureHowWrite}{}%
{[}5{]} Nature - How to write a paper, (n.d.).

\leavevmode\hypertarget{ref-rubenHowWriteScientist2012}{}%
{[}6{]} A. Ruben, 2012, 8. Am, How to Write Like a Scientist, Science
\textbar{} AAAS. (2012).

\leavevmode\hypertarget{ref-ScaryeasyWayHelp2014}{}%
{[}7{]} A scary-easy way to help you find passive voice!, A Scary-Easy
Way to Help You Find Passive Voice! \textbar{} Grammarly Blog. (2014).

\end{document}

### 2019-01-06 Paper/SupplementaryMaterial/SupplementaryMaterialsScript.R

@@ -1,0 +1,2 @@
+# Script for performing preprocessing (or other) calculations 
+# that take too long to run conveniently within an R markdown knit process.

### 2019-02-12 Paper/Manuscript.Rmd
Edits to yaml - Latex caption package is reqired for plots with gt

@@
 ---
-title: An R markdown template for Elsevier journals with examples for beginners 
+title: Article template&#58; A reproducible RMarkdown template for Elsevier's journal format with examples for beginners  
 author:
@@
     footnote: Corresponding Author
+  - name: Dana Miller
+    affiliation: CBE
   - name: Next contributor
-    affiliation: Organisation2
+    affiliation: Organization2
 address:
@@
 
-abstract: Enter the text for your abstract here
+abstract: Enter the text for your abstract here.
+All data and analysis code in this template is publicly available at [https://github.com/CenterForTheBuiltEnvironment/rmd-example](https://github.com/CenterForTheBuiltEnvironment/rmd-example) and can be freely adapted and reused.
+Suggestions or pull requests to improve this template or create additional templates are welcome.
+
 journal: "Building and Environment"
@@
 - \doublespacing
+- \usepackage{caption}
 ---

### 2019-02-12 Paper/Manuscript.Rmd
Update doc setup and load data

@@
 
-``` {r setup, echo=FALSE, cache=FALSE, results="hide", message=FALSE}
-## code chunk for libraries and global settings (e.g. applies to the whole markdown file)
-require(pander)
+``` {r libraries, echo=FALSE, cache=FALSE, results="hide", message=FALSE, warning = FALSE}
+## code chunk for libraries that are required to be installed + loaded for this template
+# note the first word after `{r...` in the chunk header above is the name of this chunk. 
+# Names must be unique and not have spaces (underscores are ok), so name with CamelCaseLikeThis or snake_case_like_this
 require(tidyverse)
-require(ggExtra)
-require(gridExtra)
+#require(ggExtra)
+#require(gridExtra)
 require(here)
-require(readxl)
-require(broom)
 require(knitr)
 require(stargazer)
-require(RColorBrewer)
-require(ggrepel)
-require(ggpmisc)
+#require(RColorBrewer)
+#require(ggrepel)
+#require(ggpmisc)
+```
 
-## numbers >= 10^5 will be denoted in scientific notation,
+```{r setup, echo=FALSE, cache=FALSE }
+# Code chunk for and global settings (e.g. applies to the whole markdown file)
+## Configures that numbers >= 10^5 will be denoted in scientific notation,
 ## and rounded to 2 digits
 options(scipen = 1, digits = 2)
+
+## Configures that by default, Rmarkdown chunks like this one won't display R messages, 
+## and won't show the code in the chunk, unless you opt-in with `echo = TRUE` in the chunk header
 knitr::opts_chunk$set(echo=FALSE, message=FALSE, fig.pos = 'h')
+```
 
+```{r themes, results = 'hide'}
 # set the gplot theme
@@
              strip.text.y = element_text(angle = 90))
-
 # 2 print and color blind safe color options from UC Berkeley's official color palette.
@@
 
-``` {r readSourceData, results="hide"}
+``` {r readSourceData}
 #read your source dataset(s) from subfolder
-# df <- read_csv(here("Paper","SupplementaryMaterial","RawData","all_test_data_and_metadata.csv"), guess_max=100000)
+df <- read_csv(here("Paper","SupplementaryMaterial","Data","iris_data.csv"), guess_max=100000, col_names = FALSE)
 ```

### 2019-02-12 Paper/Manuscript.Rmd
Add data preprocessing

@@
 ``` {r preprocessData, echo=FALSE, results="hide"}
-# here is where all the preprocessing happens
-#
+## here is where all the preprocessing happens
 # ensure you comment any sections or functions in sufficient detail that
@@
 # of the code itself
+
+# rename columns and create a new variable
+df <- df%>%
+  rename(sepal_length = X1)%>%
+  rename(sepal_width = X2) %>%
+  rename(petal_length = X3) %>%
+  rename(petal_width = X4)%>%
+  rename(iris_class = X5) %>%
+ mutate(length_ratio =  sepal_length/ petal_length)
+
+# fit linear models for relationship between sepal and petal width for each iris type
+df_lms = df %>% 
+  group_by(iris_class) %>%
+  do(fit_length = lm(sepal_length ~ petal_length, data = .))
+
+# generate convenient filtering variables
+df <- df%>%
+  mutate(
+    # identify if sepal length is more than 1 standard deviation above mean
+    long_sepal = ifelse(sepal_length>= (mean(df$sepal_length) + sd(df$sepal_length)), "Yes", "No"),
+    # identify if the petal width is a specified value, in this case exactly 1 or 2
+    integer_petal = ifelse((petal_width >= 0.9 & petal_width<= 1.1) | (petal_width>= 1.9 & petal_width<=2.1),"Yes", "No")
+  )
+
+# characterize petal length into 5 categories in increments of 1.5 cm
+df  = df%>%
+  mutate(petel_length_level = ifelse(petal_length>= 6 & petal_length <7.5,"VeryHigh",0))%>%   
+  mutate(petel_length_level = ifelse(petal_length>= 4.5 & petal_length <6,"High",petel_length_level))%>%  
+  mutate(petel_length_level = ifelse(petal_length>= 3 & petal_length <4.5,"Medium",petel_length_level))%>%
+  mutate(petel_length_level = ifelse(petal_length>= 1.5 & petal_length <3,"Low",petel_length_level))%>%
+  mutate(petel_length_level = ifelse(petal_length <1.5,"VeryLow",petel_length_level))
+
+# order new petal length categories correctly from Low to High
+df = df%>%mutate(petel_length_level = as.factor(petel_length_level))
+df$petel_length_level = factor(df$petel_length_level, levels=c("VeryLow","Low","Medium","High","VeryHigh"))
 ```
@@
 Keywords\:  
-Key 1\; Key 2\; Key 3\; Key 4\; Key 5\; Key 6 (max)
+Keyword 1\; Keyword 2\; Keyword 3\; Keyword 4\; Keyword 5\; Keyword 6 (max)
 

### 2019-02-12 Paper/Manuscript.Rmd
Add a basic table + fix variable name

@@
 #require(ggpmisc)
+require(gt)
 ```
@@
 df  = df%>%
-  mutate(petel_length_level = ifelse(petal_length>= 6 & petal_length <7.5,"VeryHigh",0))%>%   
-  mutate(petel_length_level = ifelse(petal_length>= 4.5 & petal_length <6,"High",petel_length_level))%>%  
-  mutate(petel_length_level = ifelse(petal_length>= 3 & petal_length <4.5,"Medium",petel_length_level))%>%
-  mutate(petel_length_level = ifelse(petal_length>= 1.5 & petal_length <3,"Low",petel_length_level))%>%
-  mutate(petel_length_level = ifelse(petal_length <1.5,"VeryLow",petel_length_level))
+  mutate(petal_length_level = ifelse(petal_length>= 6 & petal_length <7.5,"VeryHigh",0))%>%   
+  mutate(petal_length_level = ifelse(petal_length>= 4.5 & petal_length <6,"High",petal_length_level))%>%  
+  mutate(petal_length_level = ifelse(petal_length>= 3 & petal_length <4.5,"Medium",petal_length_level))%>%
+  mutate(petal_length_level = ifelse(petal_length>= 1.5 & petal_length <3,"Low",petal_length_level))%>%
+  mutate(petal_length_level = ifelse(petal_length <1.5,"VeryLow",petal_length_level))
 
 # order new petal length categories correctly from Low to High
-df = df%>%mutate(petel_length_level = as.factor(petel_length_level))
-df$petel_length_level = factor(df$petel_length_level, levels=c("VeryLow","Low","Medium","High","VeryHigh"))
+df = df%>%mutate(petal_length_level = as.factor(petal_length_level))
+df$petal_length_level = factor(df$petal_length_level, levels=c("VeryLow","Low","Medium","High","VeryHigh"))
 ```
@@
 -   Add your 3 -5 highlights
--   As bullets here
+-   As bulletpoints here
 -   Making sure each is less than 85 characters in length
@@
 
-# Examples
+# Methods
+
+## Markdown
+Here are some examples of how to format text with Markdown.
+
 # Heading level 1
@@
 ### Heading level 3
-Here's how to **bold** or *italic* a piece of text.
+
+Here's how to **bold** or *italicize* a piece of text.
 
@@
     + Item 2b
+
+## Tables
+Here is an example of a table created from the .csv file in the `SupplementaryMaterial/Data`folder using the `gt` package from RStudio.
+You can read more about `gt` [here](https://github.com/rstudio/gt), and there's helpful examples to further customize tables (eg with color, foototes, significant figures, re-labeling columns, and more) [here](https://github.com/allisonhorst/gt-awesome-tables).
+
+```{r}
+df %>% 
+    sample_n(5) %>%   # select 5 random rows from the dataframe
+    select(iris_class, sepal_length, petal_length, petal_length_level) %>%  # show only these columns (in this order)
+    gt () %>% #make a table
+    tab_header( 
+    title = "Example irises", # ...with this title
+    subtitle = "Data on five randomly selected irises")
+
+```
 
@@
 
-# Methods
-
 # Results

### 2019-02-12 Paper/Manuscript.Rmd
Re-order sections, add plots

@@
 #require(ggrepel)
-#require(ggpmisc)
+require(ggpmisc)
 require(gt)
@@
 
+ You may wish to have short sub-sections for:
+
+ - Terminology
+ - Objectives
+
 # Methods
@@
 
-```{r}
+```{r irisTable}
 df %>% 
@@
 ```
-
-## Citations
-Citing other literature is remarkably easy, just like this @coakleyReviewMethodsMatch2014.
-This citation key references the tag associated with an entry in Bibliography.bib (a BibTex file).
-I've found it easiest to use Zotero to manage my library of references and to generate the BibTex file, though any software that creates a valid BibTex file should work fine.
-Zotero allows you to create a 'Collection' (or folder) that gathers together all of the references used for a particular document.
-When combined with with the Better BibTex plugin, that collection can be exported to a BibTex file that is continually updated as you add or edit references in that Collection.
-Better BibTex also puts the citation key - the text after the '\@' sybmbol in the .Rmd file - on the upper right of each entry, which is convenient for adding citations.
-
-There's not much else involved in citing, as the references list gets built and formatted automatically based on the selected style.
-The only other issue I've had to look around to solve was figuring out how to combine multiple citations, which is easy when you know how.
-[@coakleyReviewMethodsMatch2014; @zhaiHumanComfortPerceived2015a]
 
@@
 
+## Citations
+Citing other literature is remarkably easy, just like this @coakleyReviewMethodsMatch2014.
+This citation key references the tag associated with an entry in Bibliography.bib (a BibTex file).
+I've found it easiest to use Zotero to manage my library of references and to generate the BibTex file, though any software that creates a valid BibTex file should work fine.
+Zotero allows you to create a 'Collection' (or folder) that gathers together all of the references used for a particular document.
+When combined with with the Better BibTex plugin, that collection can be exported to a BibTex file that is continually updated as you add or edit references in that Collection.
+Better BibTex also puts the citation key - the text after the '\@' sybmbol in the .Rmd file - on the upper right of each entry, which is convenient for adding citations.
+
+There's not much else involved in citing, as the references list gets built and formatted automatically based on the selected style.
+The only other issue I've had to look around to solve was figuring out how to combine multiple citations, which is easy when you know how.
+[@coakleyReviewMethodsMatch2014; @zhaiHumanComfortPerceived2015a]
+
+The instructions for installing BetterBibtex are [here](https://retorque.re/zotero-better-bibtex/).
+
+### Add note on citing software
+- TODO
+
 ## Other packages
-Here are lots of packages that are useful for markdown docs.
-I encourage you to search for these whenever you encounter a new thing you need to do and to propose an addition to this repository accordingly.
-For example, the Stargazer package describes linear models in a nice table.
-
-``` {r linmods, results='asis', echo=FALSE, warning=FALSE}
-lm1 = lm(hwy ~ displ + cyl + year, mpg)
-lm2 = lm(cty ~ displ + cyl + year, mpg)
-# uncomment next line for printing in console (i.e. human readable in RStudio)
-#stargazer(lm1, lm2, type="text")
-stargazer(lm1, lm2, dep.var.labels=c("Highway","City"), covariate.labels=c("Displacement","Num. cylinders", "Year"), title = "Car miles per gallon", column.sep.width = "1pt", header=FALSE)
-```
+There are lots of packages that are useful for markdown docs.
+We encourage you to search for these whenever you encounter a new thing you need to do and to propose an addition to this repository accordingly.`
 
@@
 
-# Writing style
+Here is another equation:
+
+$$ CD_{rated} = \frac{4*Q}{\pi*D^2} = 2.0~m/s$$
+
+## Writing style
 This is a little off topic for an Rmd example but a convenient place to remind our grad students about writing style.
@@
 
-## A fun way to spot passive voice:
+### A fun way to spot passive voice:
 If you can add the words ‘by zombies’ @ScaryeasyWayHelp2014 to the end of the sentence and the sentence still makes logical sense, then the sentence is in passive voice.
@@
 # Results
+Figure \@ref(fig:irises) shows petal widths by iris type.
+
+```{r irises, message=FALSE, fig.width =8, fig.height = 4.5, fig.cap = "Petal and sepal lengths from iris dataset" }
+ggplot(df, aes(x = petal_length, y = sepal_length, group = iris_class)) + 
+  geom_smooth(method=lm, se=FALSE, size = 0.5, linetype = "solid", color = "darkgray") +
+  geom_point(aes(color = iris_class, shape = long_sepal), alpha = 0.6, size=3) +
+  coord_cartesian(xlim = c(0,9), ylim = c(0,9)) + 
+  stat_poly_eq(formula = y~x, 
+               aes(color=iris_class, label = paste(..eq.label.., sep = "~~~")), 
+               parse = TRUE, label.y = c(2, 2.5, 3), 
+               label.x = 7, hjust = 1, size = 3.5) +
+  scale_y_continuous(expand=c(0,0),expression(Petal~length ~ (cm))) +
+  scale_x_continuous(expand=c(0,0),expression(Sepal~length (cm))) +
+  guides(color=guide_legend(title="Class",
+                            reverse=T, override.aes = list(shape = 15, size=3)), 
+         shape=guide_legend(title="long_sepal")) +
+  scale_color_manual (values = cal_palette4)
+```
+
+## Violin plot
+Figure \@ref(fig:petalwidths) shows petal widths by iris type.
+
+``` {r petalwidths, message=FALSE, fig.width = 6, fig.height = 4, fig.cap = "Petal widths by class" }
+df %>%
+    ggplot() + 
+    geom_violin(aes( x = iris_class, y = petal_width, color = iris_class), draw_quantiles = T, trim = F, show.legend = F) +
+    geom_boxplot(aes( x = iris_class, y = petal_width, color = iris_class), width = 0.07, show.legend = F)+
+    ggtitle("Petal widths by species") +
+    scale_color_manual(values=cal_palette4)
+
+```
 
 # Discussion
+
+# Limitations
+It is important to include a discussion of limitations in any paper. Limitations of this template include:
+
+- Figures are included after the text, not at the location where they are discussed in the text
 

### 2019-02-13 Paper/Manuscript.Rmd
Edit title

@@
 ---
-title: Article template&#58; A reproducible RMarkdown template for Elsevier's journal format with examples for beginners  
+title: A reproducible journal article with RMarkdown&#58; A  building science template with examples for beginners based on the Elsevier journal format 
 author:

### 2019-02-13 Paper/Manuscript.Rmd
Change citation format, and add software citations

@@
     base_format: rticles::elsevier_article
-csl: elsevier-with-titles.csl
+csl: advances-in-building-energy-research.csl
 geometry: "left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm"
@@
 
+```{r citationStyle, echo = FALSE}
+#The code above inside the `---` is called a YAML header.
+By default, the Elsevier template specifices a ciation format where in-text references look like this: [1], [2].
+However, this format is often less helpful when reviewing or editing , since there's no quick way of knowing what the citations refer to.
+
+#An alternative citation style is the `author-date` format, where in-text references look like this: Coakley et al., 2014; Zhai et al., 2015.
+
+# Currently this template is set up to knit with the author-date format from the journal "Advances in Building Energy Research", which is specified by the file above called `advances-in-building-energy-research.csl`.
+If you need to change the citation and bibliography style to the Elsevier template used by Elsevier journals like "Energy and Buildings", replace `advances-in-building-energy-research.csl` with `elsevier-with-titles.csl`
+```
+
 ``` {r libraries, echo=FALSE, cache=FALSE, results="hide", message=FALSE, warning = FALSE}
@@
 require(tidyverse)
-#require(ggExtra)
-#require(gridExtra)
 require(here)
@@
 require(stargazer)
-#require(RColorBrewer)
-#require(ggrepel)
 require(ggpmisc)
 require(gt)
+require(grateful)
 ```
@@
 # Introduction
-The purpose of this (very much work-in-progress) document is to provide a complete R markdown template for an Elsevier journal submission (based on the rticles repository @LaTeXJournalArticle2019), along with useful examples and packages to improve usability for folks who are just starting out with this workflow.
+The purpose of this (very much work-in-progress) document is to provide a complete R markdown template for an Elsevier journal submission (based on the rticles repository @rticles), along with useful examples and packages to improve usability for folks who are just starting out with this workflow.
 The eventual intent is to capture minimal examples of the common things that authors need to do when writing papers in R markdown; provide examples of useful packages, workflows, and tools; and provide solutions to common issues that folks encounter.
@@
 
-# Methods
+# Methods (this is a 'level 1 heading')
 
-## Markdown
+## Markdown (this is a 'level 2 heading')
 Here are some examples of how to format text with Markdown.
 
-# Heading level 1
-## Heading level 2
-### Heading level 3
+### Level 3 heading
 
@@
 
-### Add note on citing software
-- TODO
+## Other packages
+There are lots of packages that are useful for markdown docs and customizing plots.
+We encourage you to search for these whenever you encounter a new thing you need to do and to propose an addition to this repository accordingly.
+Some examples to start: `ggExtra`, `gridExtra`, `RColorBrewer`, `ggrepel`...
 
-## Other packages
-There are lots of packages that are useful for markdown docs.
-We encourage you to search for these whenever you encounter a new thing you need to do and to propose an addition to this repository accordingly.`
+### Citing software
+
+```{r include=FALSE}
+cite_packages(out.format = "md", out.dir =  here("/Paper/SupplementaryMaterial"))
+
+```
+
+It is also good practice to include citations not such for literature like journal articles and reports, but also the software used in the analysis.
+Many software packages have accompanying publications that can be cited, and even if they don't it is important to attribute which version and which software you used.
+Citing software helps to provide credit for the creators and maintainors of the software and demonstrate that people (like you!) are using the software, which is especially important for open-source tools supported through public research funding.
+
+You can use the `greatful` package to quickly generate a list of the packages for software used in an R project using the `cite_packages` function from `greatful`.
+There is an additional function called `get_citations` that generates the Bibtex-formatted citations for each package.
+It's not elegant, but you can use the first to list packages used, then generate the `pkg-refs.bib` file (saved to `/Paper/SuplementaryMaterial` by default in this template), copy the content into your overall `Bibliography.bib` file for the whole document, and then cite the packages in your document as you would another source, eg by using `@base` for an in-text citation for core R software before other packages are installed (also known as 'base R').
+
+-   base R - @base
+-   ggpmisc - @ggpmisc
+-   grateful  - @grateful
+-   gt - @gt
+-   here  - @here
+-   knitr - @knitr
+-   tidyverse -  @tidyverse
+-   rticles - @rticles
+
+```{r include = FALSE}
+pkgs <- c("tidyverse", "here", "knitr", "stargazer", "ggpmisc", "gt", "grateful", "base", "rticles")
+cites <- grateful::get_citations(pkgs, filename = "pkg-refs.bib", out.dir =  here("/Paper/SupplementaryMaterial") )
+```
 

### 2019-02-13 Paper/Manuscript.Rmd
Increase caption font size + add examples
For more options see here: https://stackoverflow.com/questions/37136731/set-caption-font-size-for-figures-in-yaml-header-of-rmarkdown

@@
 - \doublespacing
-- \usepackage{caption}
+- \usepackage[font={small}]{caption}
 ---
@@
 # Acknowledgements
-Don't forget to acknowledge the funder(s) with associated grant numbers if required. The same goes for folks who significantly assisted you with this paper but that are not authors.
+Don't forget to acknowledge the funder(s) with associated grant numbers if required.
+The same goes for folks who significantly assisted you with this paper but that are not authors.
+Eg "Agency (grant number 12345) supported this work, with cost share provided by the Center for the Built Environment.
+We thank Person1 and Person2 for assistance with data collection""
 
@@
 Descibe any relevant interests of the authors, particularly if there is a link to the research that is relatively uncommon and could be perceived as a conflict of interest.
+Otherwise, "All authors declare no conflict of interest."
 

### 2019-02-13 Paper/Manuscript.Rmd
Adjust caption size, software citation subfolder

@@
 If you need to change the citation and bibliography style to the Elsevier template used by Elsevier journals like "Energy and Buildings", replace `advances-in-building-energy-research.csl` with `elsevier-with-titles.csl`
+```
+
+```{r figureFontSize, echo = FALSE}
+# In the YAML header above, the font size for captions in this document is controlled by `[font={small}]` under `header-includes (note this is *larger* than the default size for this template*).
+For more options, see https://stackoverflow.com/questions/37136731/set-caption-font-size-for-figures-in-yaml-header-of-rmarkdown
 ```
@@
 ## Markdown (this is a 'level 2 heading')
-Here are some examples of how to format text with Markdown.
+This RMarkdown document is a special type of interactive document that can contain both code chunks (in R, python, or other languages), and text written in a format called Markdown.
+Here are some examples of how to format text with Markdown, and a [link](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf) to an RMarkdown guide.
 
@@
 
-Here's how to **bold** or *italicize* a piece of text.
-
-This is how you do a bullet point list:
+Here's how to **bold** or *italicize* a piece of text. This is how you do a bullet point list:
 
@@
 ```{r include=FALSE}
-cite_packages(out.format = "md", out.dir =  here("/Paper/SupplementaryMaterial"))
+cite_packages(out.format = "md", out.dir =  here("/Paper/SupplementaryMaterial/Software_citations"), include.rmd = TRUE, all.pkg = TRUE)
 
@@
 pkgs <- c("tidyverse", "here", "knitr", "stargazer", "ggpmisc", "gt", "grateful", "base", "rticles")
-cites <- grateful::get_citations(pkgs, filename = "pkg-refs.bib", out.dir =  here("/Paper/SupplementaryMaterial") )
+cites <- grateful::get_citations(pkgs, filename = "pkg-refs.bib", out.dir =  here("/Paper/SupplementaryMaterial/Software_citations") )
 ```
@@
 # Results
+## Scatter plot example
 Figure \@ref(fig:irises) shows petal widths by iris type.
@@
 
-## Violin plot
+## Violin plot example
 Figure \@ref(fig:petalwidths) shows petal widths by iris type.
@@
 Eg "Agency (grant number 12345) supported this work, with cost share provided by the Center for the Built Environment.
-We thank Person1 and Person2 for assistance with data collection""
+We thank Person1 and Person2 for assistance with data collection"
 
@@
 Descibe any relevant interests of the authors, particularly if there is a link to the research that is relatively uncommon and could be perceived as a conflict of interest.
-Otherwise, "All authors declare no conflict of interest."
+Otherwise : All authors declare no conflict of interest.
 

### 2019-02-13 Paper/Manuscript.Rmd
Change Figures to Images since latex produces separate directory for figures

@@
 #include a graphical abstract, without a caption
-include_graphics(here("Paper","SupplementaryMaterial","Figures","Latex_logo.png"))
+include_graphics(here("Paper","SupplementaryMaterial","Images","Latex_logo.png"))
 
@@
 ```{r correlation, fig.width = 1, fig.cap = "Correlation. Source: XKCD, xkcd.com/552"}
-include_graphics(here("Paper","SupplementaryMaterial","Figures","Correlation.png"), dpi= 500)
+include_graphics(here("Paper","SupplementaryMaterial","Images","Correlation.png"), dpi= 500)
 ```

### 2019-02-13 install.R
Try adding reqs for Binder

@@ -1,0 +1,8 @@
+install.packages("tidyverse")
+install.packages("rmarkdown")
+install.packages("ggpmisc")
+install.packages("here")
+install.packages("knitr")
+install.packages("devtools")
+remotes::install_github("rstudio/gt")
+remotes::install_github("Pakillo/grateful")

### 2019-02-13 install.R
remotes --> devtools

@@ -4,5 +4,5 @@ install.packages("ggpmisc")
 install.packages("here")
 install.packages("knitr")
 install.packages("devtools")
-remotes::install_github("rstudio/gt")
-remotes::install_github("Pakillo/grateful")
+devtools::install_github("rstudio/gt")
+devtools::install_github("Pakillo/grateful")

### 2019-02-13 Paper/Manuscript.Rmd
Update runtime date, +bookdown, rticles, -stargazer

@@
 require(knitr)
-require(stargazer)
 require(ggpmisc)

### 2019-02-13 install.R

@@ -3,6 +3,8 @@ install.packages("rmarkdown")
 install.packages("ggpmisc")
 install.packages("here")
 install.packages("knitr")
+install.packages("bookdown")
+install.packages("rticles")
 install.packages("devtools")
 devtools::install_github("rstudio/gt")
 devtools::install_github("Pakillo/grateful")

### 2019-02-13 Paper/Manuscript.Rmd
-stargazer again

@@
 ```{r include = FALSE}
-pkgs <- c("tidyverse", "here", "knitr", "stargazer", "ggpmisc", "gt", "grateful", "base", "rticles")
+pkgs <- c("tidyverse", "here", "knitr", "ggpmisc", "gt", "grateful", "base", "rticles")
 cites <- grateful::get_citations(pkgs, filename = "pkg-refs.bib", out.dir =  here("/Paper/SupplementaryMaterial/Software_citations") )

### 2019-02-13 Paper/Manuscript.Rmd
add tinytex, try specific packages instead of whole tidyverse

@@
 # Names must be unique and not have spaces (underscores are ok), so name with CamelCaseLikeThis or snake_case_like_this
-require(tidyverse)
+require(dplyr)
+require(readr)
+require(ggplot2)
+require(magrittr)
 require(here)
@@
 ```{r include = FALSE}
-pkgs <- c("tidyverse", "here", "knitr", "ggpmisc", "gt", "grateful", "base", "rticles")
+pkgs <- c("here", "knitr", "ggpmisc", "gt", "grateful", "base", "rticles")
 cites <- grateful::get_citations(pkgs, filename = "pkg-refs.bib", out.dir =  here("/Paper/SupplementaryMaterial/Software_citations") )

### 2019-02-13 install.R

@@ -1,4 +1,7 @@
-install.packages("tidyverse")
+install.packages("magrittr")
+install.packages("dplyr")
+install.packages("readr")
+install.packages("ggplot2")
 install.packages("rmarkdown")
 install.packages("ggpmisc")
 install.packages("here")
@@ -8,3 +11,5 @@ install.packages("rticles")
 install.packages("devtools")
 devtools::install_github("rstudio/gt")
 devtools::install_github("Pakillo/grateful")
+install.packages("tinytex")
+tinytex::install_tinytex()

### 2019-02-14 Paper/Manuscript.Rmd
Hyperlink in-test citations

@@
 csl: advances-in-building-energy-research.csl
+link-citations: true
+linkcolor: blue
 geometry: "left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm"

### 2019-02-14 Paper/Manuscript.Rmd
Add limitations, standardize fig size

@@
 Figure \@ref(fig:petalwidths) shows petal widths by iris type.
-
-``` {r petalwidths, message=FALSE, fig.width = 6, fig.height = 4, fig.cap = "Petal widths by class" }
+``` {r petalwidths, message=FALSE, fig.width = 8, fig.height = 4.5, fig.cap = "Iris petal widths by species" }
 df %>%
@@
     geom_boxplot(aes( x = iris_class, y = petal_width, color = iris_class), width = 0.07, show.legend = F)+
-    ggtitle("Petal widths by species") +
     scale_color_manual(values=cal_palette4)
@@
 # Limitations
-It is important to include a discussion of limitations in any paper. Limitations of this template include:
+It is important to include a discussion of limitations in any paper. Some limitations of this template include:
 
-- Figures are included after the text, not at the location where they are discussed in the text
+- Whether or not figures are included inline with text appears to depend of the version of LaTeX the document is compiled with.
+Compiling on a computer where TinyTeX is installed (eg in the Binder associated with the reporsitory for this project on GitHub) will render the figures inline, whereas compiling the same document on a Windows computer with MikTeX installed does not
+- Does not include instructions on how to align figures
 

### 2019-02-14 README.Rmd
Update to include Binder explanation + MikTeX instructions

---
title: "A reproducible journal article template in RMarkdown"
output: 
    github_document:
        toc: true # table of content true
        toc_depth: 4  # upto three depths of headings (specified by #, ## and ###)
---

#

[![Binder](http://mybinder.org/badge_logo.svg)](http://mybinder.org/v2/gh/CenterForTheBuiltEnvironment/rmd-example/master?urlpath=rstudio)

### Overview
This repository contains work in progress on an template for journal articles (eg for *Energy and Buildings*,  *Buildings and Environment*, or other journals that use the Elsevier format) along with examples for beginners. This example is slightly modified from the template from the `elsevier_article` template from the`rticles` package [here](https://github.com/rstudio/rticles).

### How to use this repo

**This repository contains all the code and data to recreate the PDF file (`Manuscript.PDF`) in the `Paper` directory, including re-creating all the figures. **

There are two ways to use this template:

1. Short-term and in the cloud (no installation required, suitable for short demonstrations or exploration, will time out after 10 minutes of inactivity)

2. Longer term and on your own computer (may require installing some additional R  packages)

#### 1. Short-term and in the cloud (no installation required)
 - Click the "launch Binder" button at the top of this page to open all of the code and data in this repository in an RStudio session where all of the required software is already installed.

- In the lower right corner, click on the `Paper` folder, click on `Manuscript.Rmd`, and press the "knit" button (on the top left). You may be prompted to update some packages the first time you knit the document - click 'yes' to accept

- When the newly-regenerated PDF document appears on the lower right version of the screen (you will see the timestamp is newer), click on it to open it. Congrats - you just re-ran this entire template!

- If you want to explore making some changes,  you can edit the `Manuscript.Rmd` file and re-knit the document to create an updated PDF.

#### 2. Longer term and on your own computer (may require installing some additional R  packages and also LaTeX if you don't have it installed already)

- Fork this repository (press the green button on the right side of this GitHub page)
- On your computer, open RStudio and create a new RStudio project using the url for your new copy of the repository, which will download all of the files locally to the computer you are working on
- Check if you have a version of LaTex installed on your computer. You can check this by typing `Sys.which(pdflatex)` in the R console. If the output is blank, you will need to install LaTeX. 
    - The recommended way to install a small version for LaTeX customized for RMarkdown is to use the TinyTex package (Mac/Windows/Linux), per the instructions [here](https://yihui.name/tinytex/)
    - If for some reason TinyTeX doesn't work for you, there is a larger LaTeX version for Windows called MikTeX. *See the more detailed instructiosn + troubleshooting for MikTeX at the bottom of this page*

- Knit the file `Manuscript.Rmd` (you can do this by navigating to it in the `Files` pane, under `/Paper`, and clicking the "knit" button on the top left of the RStudio pane)
        - You may need to install some additional error packages if prompted to do so by errors during the knit.
        - The R packages used by this template are `tidyverse`, `here`, `knitr`, `ggpmisc`, `grateful`, `gt`, `bookdown`,`rticles`, `devtools`, `rmarkdown`
            - Most of the above packages can be installed using `install_packages()`, or the "Packages" tab in RStudio. 
            - However two of the above packages have not been published to CRAN yet, and require special installation by running:

```
install.packages("devtools")
remotes::install_github("rstudio/gt")
remotes::install_github("Pakillo/grateful")

```

### Contents of this repository
The key contents are organized as follows:

- rmd-example
    - Readme.Rmd
    - .here
    - LICENSE
    - Paper
        - Bibliography.bib 
        - Manuscript.PDF
        - Manuscript.Rmd

    - SupplementaryMaterial
        - Data
        - Images
        - Software_citations
        - src

    - Manuscript_files
        - figure_latex

The top-level directory contains a`Paper` directory where most of the files for this project are stored. The `.here` file designates this location as the starting point for all file paths executed within this project. It is important to have a 'LICENSE' file to make it clear if other people have the right to reuse and build on your code, and under what conditions. See the interactive ["Choose a license"](https://choosealicense.com/) guide for help selecting one.

There are also additional files under `/Paper`:

- `Bibliography.bib` containing the BibTeX file with all the references used in this paper
- `Manuscript_files`, an automatically-created folder containing a separate .jpg image of each plot created in the manuscript
- `elsarticle.cls` for formatting the PDF 
- `elsevier-with-titles.csl` and `advances-in-building-energy-research.csl` for formatting citations with Zotero
- Some files created by the process of converting an Rmarkdown document to a PDF via LaTeX, such as `Manuscript.tex`, `Manuscript.fff`, `Manuscrpt.spl`, `Manuscript.ttt`. You do not need to edit these files.

Under `SupplementaryMaterial`:

 - `Images` contains graphics that are not created programmatically 
 - `src` is the convention for a folder that contains additional .R scripts
 - `Software_citations` contains the output of the `greatful` package for all the software cited in this template
 - `Data` is a convention for a folder containing the data used in this analysis, since this dataset is relatively small and something that can be made publicly avalible.

### Troubleshooting
- If you have knitted this document on your computer, opened the PDF (eg in Adobe Acrobat), and then try to re-knit the PDF while the file is open, the process will fail. Close the PDF, re-knit it, then re-open

#### Installing MikTex on Windows

[TinyTeX](https://yihui.name/tinytex/) is great, please try it first if you haven't already.

If you're not able to use TinyTex, then:

##### 1. Install MikTeK

https://miktex.org/download

To mamage your settings with MikTeX, open the newly-installed MikTeX Console, and 
    - Under Settings --> General, select " Always install missing packages on-the-fly"
    - Under Settings --> Directories, copy the file path for "Links to the executables have been installed in", which might look like `C:\Users\YourName\AppData\Local\Programs\MiKTeX 2.9`

If you get "pdflatex not found" error when compiling .tex file on knit, check if the environment variable set:

```{r eval=FALSE}
Sys.which("pdflatex")
pdflatex 
#Source: https://tex.stackexchange.com/questions/398218/pdflatex-not-found-windows
```

##### 2. Create and edit an .Renviron file to provide path to pdflatex.exe in the MiKTeX directory listed above
```{r eval=FALSE}
#to programmatically create an Renviron file, uncomment lines below (commented out for knit)
#install.packages("usethis")
library(usethis)
#usethis::edit_r_environ("rticles_test") # This creates an .Renviron file
```

Now, edit the .Renviron file to contain the file path to `pdflatex` (adjust for your file path as needed) and save it
```
#pdflatex
PATH=C:\\Users\\Dana\\AppData\\Local\\Programs\\MiKTeX 2.9\\miktex\\bin\\x64;"${PATH}"
```

Reference: https://community.rstudio.com/t/how-to-set-a-variable-in-renviron/5029/3
https://stackoverflow.com/questions/33650869/how-can-i-set-the-latex-path-for-sweave-in-r

##### Now try to knit the .Rmd file to a PDF!

More references:

https://stackoverflow.com/questions/48224162/knitr-wont-compile-pdf-error-in-toolsfile-path-as-absoluteoutput-file

https://medium.com/@sorenlind/create-pdf-reports-using-r-r-markdown-latex-and-knitr-on-windows-10-952b0c48bfa9

https://rickpackblog.wordpress.com/2017/10/04/miktex-install-for-r-windows-pdf-output-r-markdown/

If you experience an error like this:
"Error in tools::file_path_as_absolute(output_file)"
it may be resolved by changing the miktex settings to install required packages automatically on the fly.

Reference:  https://stackoverflow.com/questions/48224162/knitr-wont-compile-pdf-error-in-toolsfile-path-as-absoluteoutput-file

### 2019-02-15 Paper/Manuscript.Rmd
Fix figure chunk formatting, add set.seed

@@
 ```{r irisTable}
+# The next calculation involves a random draw. Setting a seed ensures the same pseudo-random selection will be made
+# each time this document is re-run with the same seed
+# Note that deliberately varying the seed and repeating pseudo-random calculations until you achieve for example  a desired but unlikely result is not ethical research practice ('seed hacking`)
+set.seed(42)
+
 df %>% 
@@
 
-```{r irises, message=FALSE, fig.width =8, fig.height = 4.5, fig.cap = "Petal and sepal lengths from iris dataset" }
+```{r irises, message=FALSE, fig.width =8, fig.height = 4, fig.align='center', fig.cap = "Petal and sepal lengths from iris dataset" }
 ggplot(df, aes(x = petal_length, y = sepal_length, group = iris_class)) + 
@@
 Figure \@ref(fig:petalwidths) shows petal widths by iris type.
-``` {r petalwidths, message=FALSE, fig.width = 8, fig.height = 4.5, fig.cap = "Iris petal widths by species" }
+
+```{r petalwidths, message=FALSE, fig.width = 5, fig.height = 3, fig.align='center', fig.cap = "Iris petal widths by species" }
 df %>%
@@
     scale_color_manual(values=cal_palette4) 
-
 ```

### 2019-02-15 Paper/Manuscript.Rmd
Add two more software citations

@@
 
-## Other packages
-There are lots of packages that are useful for markdown docs and customizing plots.
-We encourage you to search for these whenever you encounter a new thing you need to do and to propose an addition to this repository accordingly.
-Some examples to start: `ggExtra`, `gridExtra`, `RColorBrewer`, `ggrepel`...
-
 ### Citing software
@@
 -   rticles - @rticles
+-   bookdown - @bookdown
+-   rmarkdown - @rmarkdown
 
 ```{r include = FALSE}
-pkgs <- c("here", "knitr", "ggpmisc", "gt", "grateful", "base", "rticles")
+pkgs <- c("here", "knitr", "ggpmisc", "gt", "grateful", "base", "rticles", "bookdown", "rmarkdown", "tidyverse")
 cites <- grateful::get_citations(pkgs, filename = "pkg-refs.bib", out.dir =  here("/Paper/SupplementaryMaterial/Software_citations") )
@@
 $$ CD_{rated} = \frac{4*Q}{\pi*D^2} = 2.0~m/s$$
+
+## Other packages
+There are lots of packages that are useful for markdown docs and customizing plots.
+We encourage you to search for these whenever you encounter a new thing you need to do and to propose an addition to this repository accordingly.
+Some examples to start: `ggExtra`, `gridExtra`, `RColorBrewer`, `ggrepel`...
 

### 2019-02-15 Paper/Manuscript.tex
Push corresponding updated PDF

@@
 \usepackage[hyphens]{url}
-
-  \journal{Building and Environment} % Sets Journal name
-
 \usepackage{lineno} % add
@@
 \usepackage{booktabs} % book-quality tables
+%% Redefines the elsarticle footer
+%\makeatletter
+%\def\ps@pprintTitle{%
+% \let\@oddhead\@empty
+% \let\@evenhead\@empty
+% \def\@oddfoot{\it \hfill\today}%
+% \let\@evenfoot\@oddfoot}
+%\makeatother
+
+% A modified page layout
+\textwidth 6.75in
+\oddsidemargin -0.15in
+\evensidemargin -0.15in
+\textheight 9in
+\topmargin -0.5in
 %%%%%%%%%%%%%%%% end my additions to header
@@
             pdfauthor={},
-            pdftitle={An R markdown template for Elsevier journals with examples for beginners},
-            colorlinks=false,
+            pdftitle={A reproducible journal article with RMarkdown: A building science template with examples for beginners based on the Elsevier journal format},
+            colorlinks=true,
             urlcolor=blue,
-            linkcolor=magenta,
+            linkcolor=blue,
             pdfborder={0 0 0}}
 \urlstyle{same}  % don't use monospace font for urls
-
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
 \setcounter{secnumdepth}{5}
+\usepackage{setspace}
+\doublespacing
+\usepackage[font={small}]{caption}
 % Pandoc toggle for numbering sections (defaults to be off)
@@
 \doublespacing
+\usepackage[font={small}]{caption}
 
-\usepackage{amsthm}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{proposition}{Proposition}[section]
-\theoremstyle{definition}
-\newtheorem{example}{Example}[section]
-\theoremstyle{definition}
-\newtheorem{exercise}{Exercise}[section]
-\theoremstyle{remark}
-\newtheorem*{remark}{Remark}
-\newtheorem*{solution}{Solution}
+\usepackage[nomarkers]{endfloat}
+
 \begin{document}
@@
 
-  \title{An R markdown template for Elsevier journals with examples for beginners}
+  \title{A reproducible journal article with RMarkdown: A building science
+template with examples for beginners based on the Elsevier journal
+format}
     \author[CBE]{Paul Raftery\corref{c1}}
@@
    \cortext[c1]{Corresponding Author}
-    \author[Organisation2]{Next contributor}
+    \author[CBE]{Dana Miller}
+
+    \author[Organization2]{Next contributor}
 
@@
   \begin{abstract}
-  Enter the text for your abstract here
+  Enter the text for your abstract here. All data and analysis code in
+  this template is publicly available at
+  \url{https://github.com/CenterForTheBuiltEnvironment/rmd-example} and
+  can be freely adapted and reused. Suggestions or pull requests to
+  improve this template or create additional templates are welcome.
   \end{abstract}
@@
 Keywords:\\
-Key 1; Key 2; Key 3; Key 4; Key 5; Key 6 (max)
+Keyword 1; Keyword 2; Keyword 3; Keyword 4; Keyword 5; Keyword 6 (max)
 
@@
 \item
-  As bullets here
+  As bulletpoints here
 \item
@@
 \textbf{Graphical Abstract}\\
-\includegraphics{C:/Users/praft/Documents/GitHub/rmd-example/Paper/SupplementaryMaterial/Figures/Latex_logo.png}
+\includegraphics{C:/Users/Dana/3_Professional/CBE/Reproducibility/cbe-rmd-example/Paper/SupplementaryMaterial/Images/Latex_logo.png}
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 a complete R markdown template for an Elsevier journal submission (based
-on the rticles repository {[}1{]}), along with useful examples and
+on the rticles repository Allaire et al.
+(\protect\hyperlink{ref-rticles}{2017})), along with useful examples and
 packages to improve usability for folks who are just starting out with
@@
 
-\hypertarget{examples}{%
-\section{Examples}\label{examples}}
+You may wish to have short sub-sections for:
 
-\hypertarget{heading-level-1}{%
-\section{Heading level 1}\label{heading-level-1}}
+\begin{itemize}
+\tightlist
+\item
+  Terminology
+\item
+  Objectives
+\end{itemize}
 
-\hypertarget{heading-level-2}{%
-\subsection{Heading level 2}\label{heading-level-2}}
+\section{\texorpdfstring{Methods (this is a `level 1
+heading')}{Methods (this is a level 1 heading)}}\label{methods-this-is-a-level-1-heading}
 
-\hypertarget{heading-level-3}{%
-\subsubsection{Heading level 3}\label{heading-level-3}}
+\subsection{\texorpdfstring{Markdown (this is a `level 2
+heading')}{Markdown (this is a level 2 heading)}}\label{markdown-this-is-a-level-2-heading}
 
-Here's how to \textbf{bold} or \emph{italic} a piece of text.
+This RMarkdown document is a special type of interactive document that
+can contain both code chunks (in R, python, or other languages), and
+text written in a format called Markdown. Here are some examples of how
+to format text with Markdown, and a
+\href{https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf}{link}
+to an RMarkdown guide.
 
-This is how you do a bullet point list:
+\subsubsection{Level 3 heading}\label{level-3-heading}
+
+Here's how to \textbf{bold} or \emph{italicize} a piece of text. This is
+how you do a bullet point list:
 
@@
 
-\hypertarget{citations}{%
-\subsection{Citations}\label{citations}}
+\subsection{Tables}\label{tables}
 
-Citing other literature is remarkably easy, just like this {[}2{]}. This
+Here is an example of a table created from the .csv file in the
+\texttt{SupplementaryMaterial/Data}folder using the \texttt{gt} package
+from RStudio. You can read more about \texttt{gt}
+\href{https://github.com/rstudio/gt}{here}, and there's helpful examples
+to further customize tables (eg with color, foototes, significant
+figures, re-labeling columns, and more)
+\href{https://github.com/allisonhorst/gt-awesome-tables}{here}.
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+
+\begin{longtable}{lrrc}
+\caption*{
+\large Example irises\\ 
+\small Data on five randomly selected irises\\ 
+} \\ 
+\toprule
+iris\_class & sepal\_length & petal\_length & petal\_length\_level \\ 
+\midrule
+Iris-virginica & 6.4 & 5.5 & High \\ 
+Iris-virginica & 6.9 & 5.4 & High \\ 
+Iris-setosa & 4.4 & 1.3 & VeryLow \\ 
+Iris-virginica & 7.7 & 6.7 & VeryHigh \\ 
+Iris-versicolor & 5.0 & 3.3 & Medium \\ 
+\bottomrule
+\end{longtable}
+
+\subsection{Cross-referencing}\label{cross-referencing}
+
+This is how you refer to a figure in your text: Figure
+\ref{fig:correlation}. Simply reference the title of the code chunk, and
+ensure that the code chunk includes a figure caption.
+
+\begin{figure}
+\centering
+\includegraphics{C:/Users/Dana/3_Professional/CBE/Reproducibility/cbe-rmd-example/Paper/SupplementaryMaterial/Images/Correlation.png}
+\caption{\label{fig:correlation}Correlation. Source: XKCD, xkcd.com/552}
+\end{figure}
+
+\subsection{Calculations in text}\label{calculations-in-text}
+
+The holy grail of markdown - doing all of your calculations in the same
+file so you never need to worry about updating them after
+someone\footnote{Often I'm the someone, sorry CBE grad students. Also,
+  look, it's an example of a footnote!} asks you to make
+changes\ldots{}. again! It's as easy as pi: 3.14. Incidentally,you can
+selectively override the `global' options set at the beginning, to say
+for example, show more decimals: 3.1416.
+
+This is an example of outputing the result of a calculation that you
+perform within a code chunk in the document somewhere prior to the
+location where you first refer to it: 36.
+
+\subsection{Citations}\label{citations}
+
+Citing other literature is remarkably easy, just like this Coakley,
+Raftery, \& Keane
+(\protect\hyperlink{ref-coakleyReviewMethodsMatch2014}{2014}). This
 citation key references the tag associated with an entry in
@@
 other issue I've had to look around to solve was figuring out how to
-combine multiple citations, which is easy when you know how. {[}2,3{]}
+combine multiple citations, which is easy when you know how. (Coakley et
+al., \protect\hyperlink{ref-coakleyReviewMethodsMatch2014}{2014}; Zhai
+et al., \protect\hyperlink{ref-zhaiHumanComfortPerceived2015a}{2015})
 
-\hypertarget{cross-referencing}{%
-\subsection{Cross-referencing}\label{cross-referencing}}
+The instructions for installing BetterBibtex are
+\href{https://retorque.re/zotero-better-bibtex/}{here}.
 
-This is how you refer to a figure in your text: Figure
-\ref{fig:correlation}. Simply reference the title of the code chunk, and
-ensure that the code chunk includes a figure caption.
+\subsubsection{Citing software}\label{citing-software}
 
-\begin{figure}
-\centering
-\includegraphics{C:/Users/praft/Documents/GitHub/rmd-example/Paper/SupplementaryMaterial/Figures/Correlation.png}
-\caption{\label{fig:correlation}Correlation. Source: XKCD, xkcd.com/552}
-\end{figure}
+It is also good practice to include citations not such for literature
+like journal articles and reports, but also the software used in the
+analysis. Many software packages have accompanying publications that can
+be cited, and even if they don't it is important to attribute which
+version and which software you used. Citing software helps to provide
+credit for the creators and maintainors of the software and demonstrate
+that people (like you!) are using the software, which is especially
+important for open-source tools supported through public research
+funding.
 
-\hypertarget{calculations-in-text}{%
-\subsection{Calculations in text}\label{calculations-in-text}}
+You can use the \texttt{greatful} package to quickly generate a list of
+the packages for software used in an R project using the
+\texttt{cite\_packages} function from \texttt{greatful}. There is an
+additional function called \texttt{get\_citations} that generates the
+Bibtex-formatted citations for each package. It's not elegant, but you
+can use the first to list packages used, then generate the
+\texttt{pkg-refs.bib} file (saved to
+\texttt{/Paper/SuplementaryMaterial} by default in this template), copy
+the content into your overall \texttt{Bibliography.bib} file for the
+whole document, and then cite the packages in your document as you would
+another source, eg by using \texttt{@base} for an in-text citation for
+core R software before other packages are installed (also known as `base
+R').
 
-The holy grail of markdown - doing all of your calculations in the same
-file so you never need to worry about updating them after
-someone\footnote{Often I'm the someone, sorry CBE grad students. Also,
-  look, it's an example of a footnote!} asks you to make
-changes\ldots{}. again! It's as easy as pi: 3.14. Incidentally,you can
-selectively override the `global' options set at the beginning, to say
-for example, show more decimals: 3.1416.
+\begin{itemize}
+\tightlist
+\item
+  base R - R Core Team (\protect\hyperlink{ref-base}{2018})
+\item
+  ggpmisc - Aphalo (\protect\hyperlink{ref-ggpmisc}{2016})
+\item
+  grateful - Rodriguez-Sanchez (\protect\hyperlink{ref-grateful}{2018})
+\item
+  gt - Iannone, Cheng, \& Schloerke (\protect\hyperlink{ref-gt}{2019})
+\item
+  here - Müller (\protect\hyperlink{ref-here}{2017})
+\item
+  knitr - Xie (\protect\hyperlink{ref-knitr}{2018})
+\item
+  tidyverse - Wickham (\protect\hyperlink{ref-tidyverse}{2017})
+\item
+  rticles - Allaire et al. (\protect\hyperlink{ref-rticles}{2017})
+\item
+  bookdown - Xie (\protect\hyperlink{ref-bookdown}{2016})
+\item
+  rmarkdown - Allaire et al. (\protect\hyperlink{ref-rmarkdown}{2018})
+\end{itemize}
 
-This is an example of outputing the result of a calculation that you
-perform within a code chunk in the document somewhere prior to the
-location where you first refer to it: 36.
-
-\hypertarget{other-packages}{%
-\subsection{Other packages}\label{other-packages}}
-
-Here are lots of packages that are useful for markdown docs. I encourage
-you to search for these whenever you encounter a new thing you need to
-do and to propose an addition to this repository accordingly. For
-example, the Stargazer package describes linear models in a nice table.
-
-\begin{table}[!htbp] \centering 
-  \caption{Car miles per gallon} 
-  \label{} 
-\begin{tabular}{@{\extracolsep{1pt}}lcc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{2}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-3} 
-\\[-1.8ex] & Highway & City \\ 
-\\[-1.8ex] & (1) & (2)\\ 
-\hline \\[-1.8ex] 
- Displacement & $-$2.100$^{***}$ & $-$1.300$^{***}$ \\ 
-  & (0.520) & (0.340) \\ 
-  & & \\ 
- Num. cylinders & $-$1.300$^{***}$ & $-$1.200$^{***}$ \\ 
-  & (0.410) & (0.270) \\ 
-  & & \\ 
- Year & 0.150$^{***}$ & 0.071$^{**}$ \\ 
-  & (0.055) & (0.036) \\ 
-  & & \\ 
- Constant & $-$259.000$^{**}$ & $-$114.000 \\ 
-  & (109.000) & (72.000) \\ 
-  & & \\ 
-\hline \\[-1.8ex] 
-Observations & 234 & 234 \\ 
-R$^{2}$ & 0.620 & 0.670 \\ 
-Adjusted R$^{2}$ & 0.610 & 0.670 \\ 
-Residual Std. Error (df = 230) & 3.700 & 2.500 \\ 
-F Statistic (df = 3; 230) & 124.000$^{***}$ & 158.000$^{***}$ \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
-\end{tabular} 
-\end{table}
-
-\hypertarget{equations-and-math}{%
-\subsection{Equations and math}\label{equations-and-math}}
+\subsection{Equations and math}\label{equations-and-math}
 
@@
 
-\hypertarget{writing-style}{%
-\section{Writing style}\label{writing-style}}
+Here is another equation:
+
+\[ CD_{rated} = \frac{4*Q}{\pi*D^2} = 2.0~m/s\]
+
+\subsection{Other packages}\label{other-packages}
+
+There are lots of packages that are useful for markdown docs and
+customizing plots. We encourage you to search for these whenever you
+encounter a new thing you need to do and to propose an addition to this
+repository accordingly. Some examples to start: \texttt{ggExtra},
+\texttt{gridExtra}, \texttt{RColorBrewer}, \texttt{ggrepel}\ldots{}
+
+\subsection{Writing style}\label{writing-style}
 
@@
 that information is more accurately reported by authors when writing in
-active voice. For example, research {[}4{]} has shown that the ``active
-{[}voice{]} offers a neutral structure for conveying information''.
-Authorship guides for highly regarded journals often indicate a
-preference for the active voice instead of passive:
+active voice. For example, research Klenbort \& Anisfeld
+(\protect\hyperlink{ref-klenbortMarkednessPerspectiveInterpretation1974}{1974})
+has shown that the ``active {[}voice{]} offers a neutral structure for
+conveying information''. Authorship guides for highly regarded journals
+often indicate a preference for the active voice instead of passive:
 
@@
   readers find concepts and results to be conveyed more clearly if
-  written directly.''{[}5{]}
+  written directly.''(``Nature - How to write a paper,'' n.d.)
 \item
@@
   a lZap library \ldots{},' not `To address this possibility, a lZap
-  library was constructed\ldots{}').'' {[}6{]}
+  library was constructed\ldots{}').'' Ruben, 2012, \& Am
+  (\protect\hyperlink{ref-rubenHowWriteScientist2012}{2012})
 \end{itemize}
@@
 
-\hypertarget{a-fun-way-to-spot-passive-voice}{%
-\subsection{A fun way to spot passive
-voice:}\label{a-fun-way-to-spot-passive-voice}}
+\subsubsection{A fun way to spot passive
+voice:}\label{a-fun-way-to-spot-passive-voice}
 
-If you can add the words `by zombies' {[}7{]} to the end of the sentence
-and the sentence still makes logical sense, then the sentence is in
-passive voice. You can also switch on the grammar settings in Microsoft
-Word's spelling and grammar checker and it will show up that way.
+If you can add the words `by zombies' (``A scary-easy way to help you
+find passive voice!''
+\protect\hyperlink{ref-ScaryeasyWayHelp2014}{2014}) to the end of the
+sentence and the sentence still makes logical sense, then the sentence
+is in passive voice. You can also switch on the grammar settings in
+Microsoft Word's spelling and grammar checker and it will show up that
+way.
 
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\section{Results}\label{results}
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\subsection{Scatter plot example}\label{scatter-plot-example}
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+Figure \ref{fig:irises} shows petal widths by iris type.
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\begin{figure}[h]
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+{\centering \includegraphics{Manuscript_files/figure-latex/irises-1}
+
+}
+
+\caption{Petal and sepal lengths from iris dataset}\label{fig:irises}
+\end{figure}
+
+\subsection{Violin plot example}\label{violin-plot-example}
+
+Figure \ref{fig:petalwidths} shows petal widths by iris type.
+
+\begin{figure}[h]
+
+{\centering \includegraphics{Manuscript_files/figure-latex/petalwidths-1}
+
+}
+
+\caption{Iris petal widths by species}\label{fig:petalwidths}
+\end{figure}
+
+\section{Discussion}\label{discussion}
+
+\section{Limitations}\label{limitations}
+
+It is important to include a discussion of limitations in any paper.
+Some limitations of this template include:
+
+\begin{itemize}
+\tightlist
+\item
+  Whether or not figures are included inline with text appears to depend
+  of the version of LaTeX the document is compiled with. Compiling on a
+  computer where TinyTeX is installed (eg in the Binder associated with
+  the reporsitory for this project on GitHub) will render the figures
+  inline, whereas compiling the same document on a Windows computer with
+  MikTeX installed does not
+\item
+  Does not include instructions on how to align figures
+\end{itemize}
+
+\section{Conclusion}\label{conclusion}
+
+\section{Acknowledgements}\label{acknowledgements}
 
@@
 if required. The same goes for folks who significantly assisted you with
-this paper but that are not authors.
+this paper but that are not authors. Eg ``Agency (grant number 12345)
+supported this work, with cost share provided by the Center for the
+Built Environment. We thank Person1 and Person2 for assistance with data
+collection''
 
-\hypertarget{declaration-of-interest}{%
-\section{Declaration of interest}\label{declaration-of-interest}}
+\section{Declaration of interest}\label{declaration-of-interest}
 
@@
 a link to the research that is relatively uncommon and could be
-perceived as a conflict of interest.
+perceived as a conflict of interest. Otherwise : All authors declare no
+conflict of interest.
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-LaTeXJournalArticle2019}{}%
-{[}1{]} LaTeX Journal Article Templates for R Markdown. Contribute to
-rstudio/rticles development by creating an account on GitHub, (2019).
+\hypertarget{ref-ScaryeasyWayHelp2014}{}
+A scary-easy way to help you find passive voice! (2014). \emph{A
+scary-easy way to help you find passive voice! \textbar{} Grammarly
+Blog}.
+https://www.grammarly.com/blog/a-scary-easy-way-to-help-you-find-passive-voice/.
 
-\leavevmode\hypertarget{ref-coakleyReviewMethodsMatch2014}{}%
-{[}2{]} D. Coakley, P. Raftery, M. Keane, A review of methods to match
-building energy simulation models to measured data, Renewable and
-Sustainable Energy Reviews. 37 (2014) 123--141.
-doi:\href{https://doi.org/10.1016/j.rser.2014.05.007}{10.1016/j.rser.2014.05.007}.
+\hypertarget{ref-rticles}{}
+Allaire, J., R Foundation, Wickham, H., Journal of Statistical Software,
+Xie, Y., Vaidyanathan, R., \ldots{} Yu, M. (2017). \emph{Rticles:
+Article formats for r markdown}. Retrieved from
+\url{https://CRAN.R-project.org/package=rticles}
 
-\leavevmode\hypertarget{ref-zhaiHumanComfortPerceived2015a}{}%
-{[}3{]} Y. Zhai, Y. Zhang, H. Zhang, W. Pasut, E. Arens, Q. Meng, Human
-comfort and perceived air quality in warm and humid environments with
-ceiling fans, Building and Environment. 90 (2015) 178--185.
-doi:\href{https://doi.org/10.1016/j.buildenv.2015.04.003}{10.1016/j.buildenv.2015.04.003}.
+\hypertarget{ref-rmarkdown}{}
+Allaire, J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
+A., \ldots{} Chang, W. (2018). \emph{Rmarkdown: Dynamic documents for
+r}. Retrieved from \url{https://CRAN.R-project.org/package=rmarkdown}
 
-\leavevmode\hypertarget{ref-klenbortMarkednessPerspectiveInterpretation1974}{}%
-{[}4{]} I. Klenbort, M. Anisfeld, Markedness and perspective in the
-interpretation of the active and passive voice, Quarterly Journal of
-Experimental Psychology. 26 (1974) 189--195.
-doi:\href{https://doi.org/10.1080/14640747408400404}{10.1080/14640747408400404}.
+\hypertarget{ref-ggpmisc}{}
+Aphalo, P. J. (2016). \emph{Learn r ...as you learnt your mother
+tongue}. Leanpub. Retrieved from \url{https://leanpub.com/learnr}
 
-\leavevmode\hypertarget{ref-NatureHowWrite}{}%
-{[}5{]} Nature - How to write a paper, (n.d.).
+\hypertarget{ref-coakleyReviewMethodsMatch2014}{}
+Coakley, D., Raftery, P., \& Keane, M. (2014). A review of methods to
+match building energy simulation models to measured data.
+\emph{Renewable and Sustainable Energy Reviews}, \emph{37}(Supplement
+C), 123--141. \url{https://doi.org/10.1016/j.rser.2014.05.007}
 
-\leavevmode\hypertarget{ref-rubenHowWriteScientist2012}{}%
-{[}6{]} A. Ruben, 2012, 8. Am, How to Write Like a Scientist, Science
-\textbar{} AAAS. (2012).
+\hypertarget{ref-gt}{}
+Iannone, R., Cheng, J., \& Schloerke, B. (2019). \emph{Gt: Easily create
+presentation-ready display tables}. Retrieved from
+\url{https://github.com/rstudio/gt}
 
-\leavevmode\hypertarget{ref-ScaryeasyWayHelp2014}{}%
-{[}7{]} A scary-easy way to help you find passive voice!, A Scary-Easy
-Way to Help You Find Passive Voice! \textbar{} Grammarly Blog. (2014).
+\hypertarget{ref-klenbortMarkednessPerspectiveInterpretation1974}{}
+Klenbort, I., \& Anisfeld, M. (1974). Markedness and perspective in the
+interpretation of the active and passive voice. \emph{Quarterly Journal
+of Experimental Psychology}, \emph{26}(2), 189--195.
+\url{https://doi.org/10.1080/14640747408400404}
+
+\hypertarget{ref-here}{}
+Müller, K. (2017). \emph{Here: A simpler way to find your files}.
+Retrieved from \url{https://CRAN.R-project.org/package=here}
+
+\hypertarget{ref-NatureHowWrite}{}
+Nature - How to write a paper. (n.d.).
+https://www.nature.com/authors/author\_resources/how\_write.html.
+
+\hypertarget{ref-base}{}
+R Core Team. (2018). \emph{R: A language and environment for statistical
+computing}. Vienna, Austria: R Foundation for Statistical Computing.
+Retrieved from \url{https://www.R-project.org/}
+
+\hypertarget{ref-grateful}{}
+Rodriguez-Sanchez, F. (2018). \emph{Grateful: Facilitate citation of r
+packages}. Retrieved from \url{https://github.com/Pakillo/grateful}
+
+\hypertarget{ref-rubenHowWriteScientist2012}{}
+Ruben, A., 2012, \& Am, 8. (2012). How to Write Like a Scientist.
+\emph{Science \textbar{} AAAS}.
+https://www.sciencemag.org/careers/2012/03/how-write-scientist.
+
+\hypertarget{ref-tidyverse}{}
+Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
+'tidyverse'}. Retrieved from
+\url{https://CRAN.R-project.org/package=tidyverse}
+
+\hypertarget{ref-bookdown}{}
+Xie, Y. (2016). \emph{Bookdown: Authoring books and technical documents
+with R markdown}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from
+\url{https://github.com/rstudio/bookdown}
+
+\hypertarget{ref-knitr}{}
+Xie, Y. (2018). \emph{Knitr: A general-purpose package for dynamic
+report generation in r}. Retrieved from \url{https://yihui.name/knitr/}
+
+\hypertarget{ref-zhaiHumanComfortPerceived2015a}{}
+Zhai, Y., Zhang, Y., Zhang, H., Pasut, W., Arens, E., \& Meng, Q.
+(2015). Human comfort and perceived air quality in warm and humid
+environments with ceiling fans. \emph{Building and Environment},
+\emph{90}, 178--185.
+\url{https://doi.org/10.1016/j.buildenv.2015.04.003}
 

### 2019-02-26 Paper/Manuscript.Rmd
Add step-by-step citation instructions + some spelling fixes

--- 2019-02-15 Paper/Manuscript.Rmd
@@
 # Introduction
-The purpose of this (very much work-in-progress) document is to provide a complete R markdown template for an Elsevier journal submission (based on the rticles repository @rticles), along with useful examples and packages to improve usability for folks who are just starting out with this workflow.
+The purpose of this (very much work-in-progress) document is to provide a complete R markdown template for an Elsevier journal submission (based on the [rticles](https://github.com/rstudio/rticles) repository @rticles), along with useful examples and packages to improve usability for folks who are just starting out with this workflow.
 The eventual intent is to capture minimal examples of the common things that authors need to do when writing papers in R markdown; provide examples of useful packages, workflows, and tools; and provide solutions to common issues that folks encounter.
@@
 Here is an example of a table created from the .csv file in the `SupplementaryMaterial/Data`folder using the `gt` package from RStudio.
-You can read more about `gt` [here](https://github.com/rstudio/gt), and there's helpful examples to further customize tables (eg with color, foototes, significant figures, re-labeling columns, and more) [here](https://github.com/allisonhorst/gt-awesome-tables).
+You can read more about `gt` [here](https://github.com/rstudio/gt), and there's helpful examples to further customize tables (eg with color, footnotes, significant figures, re-labeling columns, and more) [here](https://github.com/allisonhorst/gt-awesome-tables).
 
@@
 ## Citations
+### Citing literature
 Citing other literature is remarkably easy, just like this @coakleyReviewMethodsMatch2014.
@@
 When combined with with the Better BibTex plugin, that collection can be exported to a BibTex file that is continually updated as you add or edit references in that Collection.
-Better BibTex also puts the citation key - the text after the '\@' sybmbol in the .Rmd file - on the upper right of each entry, which is convenient for adding citations.
+Better BibTex also puts the citation key - the text after the '\@' symbol in the .Rmd file - on the upper right of each entry, which is convenient for adding citations.
 
@@
 
-The instructions for installing BetterBibtex are [here](https://retorque.re/zotero-better-bibtex/).
+**Step-by-step instructions for creating a new Zotero collection and adding a citation**
+1. Install BetterBibtex with the instructions [here](https://retorque.re/zotero-better-bibtex/).
 
-## Other packages
-There are lots of packages that are useful for markdown docs and customizing plots.
-We encourage you to search for these whenever you encounter a new thing you need to do and to propose an addition to this repository accordingly.
-Some examples to start: `ggExtra`, `gridExtra`, `RColorBrewer`, `ggrepel`...
+2.
+First, create a new Zotero collection for this paper in your own Zotero library.
+You can add existing references by dragging and dropping them into the folder for this collection, or add new references from the web with this collection highlighted.
+
+3.
+In the far left Zotero pane, right-click on the folder for this collection, select "Export Collection", and selection "Better BibTeX" and the format, and tick the "Keep updated" box (this will automatically update the stored .bib file when you add new items to this collection).
+Save this .bib file in a location together with the document you are working on (in this repository, it's under `/Paper`)
+    - Note - the dynamic updating *only* works if you export the whole collection to one .bib file, so clicking on the items in the collection and exporting them individually or together will not work
+
+4.
+To add a citation in the text, first make sure that the item is saved in the Zotero collection that is linked to the automatically-updating .bib file for your document.
+Then click on that item in Zotero and copy the citation key in the right-hand pane (eg *schiavonThermalComfortPerceived2017*)
+
+5.
+In your document, paste the citation key preceded by an @ sign, like this: `@schiavonThermalComfortPerceived2017` in the text where you want the citation to appear.
+If you want it wrapped in parentheses, do this `[@schiavonThermalComfortPerceived2017]`
 
@@
 cite_packages(out.format = "md", out.dir =  here("/Paper/SupplementaryMaterial/Software_citations"), include.rmd = TRUE, all.pkg = TRUE)
-
 ```
 
-It is also good practice to include citations not such for literature like journal articles and reports, but also the software used in the analysis.
-Many software packages have accompanying publications that can be cited, and even if they don't it is important to attribute which version and which software you used.
-Citing software helps to provide credit for the creators and maintainors of the software and demonstrate that people (like you!) are using the software, which is especially important for open-source tools supported through public research funding.
+It is also good practice to include citations for the software used in your analysis.
+After all, the software is also one of the tools you used to carry out your research.
+It is important to attribute which software and which versions you used you used so other people can better understand your methods.
+There have been cases where running the same analysis code with different versions of the same software produces different results!
+In addition, citing software helps to provide credit for the creators and maintainers of the software and demonstrate that people are using it, which is especially important for open-source tools supported through public research funding.
+Just like software often includes a `LICENSE` file, a good practice is for citation instructions to be included in a `CITATION` file associated with the project.
+Many software packages have accompanying publications that can be cited, and if they don't there is usually a project website or code repository.
 
-You can use the `greatful` package to quickly generate a list of the packages for software used in an R project using the `cite_packages` function from `greatful`.
-There is an additional function called `get_citations` that generates the Bibtex-formatted citations for each package.
-It's not elegant, but you can use the first to list packages used, then generate the `pkg-refs.bib` file (saved to `/Paper/SuplementaryMaterial` by default in this template), copy the content into your overall `Bibliography.bib` file for the whole document, and then cite the packages in your document as you would another source, eg by using `@base` for an in-text citation for core R software before other packages are installed (also known as 'base R').
+**Instructions for adding software citations with `grateful` and Zotero**
 
--   base R - @base
--   ggpmisc - @ggpmisc
--   grateful  - @grateful
--   gt - @gt
--   here  - @here
--   knitr - @knitr
--   tidyverse -  @tidyverse
--   rticles - @rticles
+```{r echo=TRUE, message=FALSE, warning=FALSE}
+# Example of generating BibTex file for the packages listed below
+# Note the output location is specified 
+pkgs <- c("here", "knitr", "ggpmisc", "gt", "grateful", "base", "rticles", "bookdown", 
+          "rmarkdown", "tidyverse")
+cites <- grateful::get_citations(pkgs, filename = "pkg-refs.bib",
+                    out.dir =  here("/Paper/SupplementaryMaterial/Software_citations") )
+```
 
-```{r include = FALSE}
-pkgs <- c("here", "knitr", "ggpmisc", "gt", "grateful", "base", "rticles")
-cites <- grateful::get_citations(pkgs, filename = "pkg-refs.bib", out.dir =  here("/Paper/SupplementaryMaterial/Software_citations") )
-```
+5.
+To add software citations, you can use the `greatful` package's `get_citations` function to generate the BibTeX-formatted citation for each package.
+Note that all of the packages have to be loaded into your working session, and listed before calling `get_citations` for this to work.
+
+6.
+Next, open the `pkg-refs.bib` file generated by `greatful` in a text editor like Notepad, highlight and copy the text, and then in Zotero with the collection for this project highlighted,  click `File` --> `Import from clipboard`.
+Now all the software references have been added to the .bib file that will continue to be updated for this project!
+
+7. Cite the software somewhere in your methods section. Here's an example:
+
+>This paper used the free and open source R statistical computing language [@base] with `tidyverse` [@tidyverse] software for all analysis, with the additional software packages `ggpmisc` [@ggpmisc] for graphics, `gt` [@gt]for tables, `here` [@here] for file path management, `grateful` [@grateful] for software citation, `rmarkdown` [@rmarkdown] for interactive notebooks, and `knitr` [@knitr], `rticles` [@rticles], and `bookdown` [@bookdown] to create a journal-formatted PDF and other markdown and html files.
 
@@
 - Whether or not figures are included inline with text appears to depend of the version of LaTeX the document is compiled with.
-Compiling on a computer where TinyTeX is installed (eg in the Binder associated with the reporsitory for this project on GitHub) will render the figures inline, whereas compiling the same document on a Windows computer with MikTeX installed does not
+Compiling on a computer where TinyTeX is installed (eg in the Binder associated with the repository for this project on GitHub) will render the figures inline, whereas compiling the same document on a Windows computer with MikTeX installed does not
 - Does not include instructions on how to align figures
@@
 # Declaration of interest
-Descibe any relevant interests of the authors, particularly if there is a link to the research that is relatively uncommon and could be perceived as a conflict of interest.
+Describe any relevant interests of the authors, particularly if there is a link to the research that is relatively uncommon and could be perceived as a conflict of interest.
 Otherwise : All authors declare no conflict of interest.

### 2019-02-27 Paper/Manuscript.Rmd
Add note + spellcheck example

@@
 pkgs <- c("here", "knitr", "ggpmisc", "gt", "grateful", "base", "rticles", "bookdown", 
-          "rmarkdown", "tidyverse")
+          "rmarkdown")
 cites <- grateful::get_citations(pkgs, filename = "pkg-refs.bib",

### 2019-02-27 Paper/Manuscript.Rmd
Add grateful note + spellcheck info

@@
 
+## Spellcheck
+If you are using RStudio, press F7 or go to `Edit` --> `Check spelling` in the menu.
+
 ## Cross-referencing
@@
 # Note the output location is specified 
+# Note tidyverse is not included in the list below because it's a large package to install on Binder, if you are running this on your own computer I recommend including it if you want to cite it
 pkgs <- c("here", "knitr", "ggpmisc", "gt", "grateful", "base", "rticles", "bookdown", 

### 2020-06-12 Paper/Manuscript.tex
Re-knit w/ fixing name of graphic

@@
 \usepackage[hyphens]{url}
+
+  \journal{Building and Environment} % Sets Journal name
+
 \usepackage{lineno} % add
@@
 
-\bibliographystyle{elsarticle-harv}
-\biboptions{sort&compress} % For natbib
 \usepackage{graphicx}
 \usepackage{booktabs} % book-quality tables
-%% Redefines the elsarticle footer
-%\makeatletter
-%\def\ps@pprintTitle{%
-% \let\@oddhead\@empty
-% \let\@evenhead\@empty
-% \def\@oddfoot{\it \hfill\today}%
-% \let\@evenfoot\@oddfoot}
-%\makeatother
-
-% A modified page layout
-\textwidth 6.75in
-\oddsidemargin -0.15in
-\evensidemargin -0.15in
-\textheight 9in
-\topmargin -0.5in
 %%%%%%%%%%%%%%%% end my additions to header
@@
 \usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
+\bibliographystyle{elsarticle-harv}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{longtable}
-\usepackage{graphicx}
-% We will generate all images so they have a width \maxwidth. This means
-% that they will get their normal width if they fit onto the page, but
-% are scaled down if they would overflow the margins.
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
-\else\Gin@nat@width\fi}
-\makeatother
-\let\Oldincludegraphics\includegraphics
-\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
 \ifxetex
@@
 \urlstyle{same}  % don't use monospace font for urls
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
+
 \setcounter{secnumdepth}{5}
-\usepackage{setspace}
-\doublespacing
-\usepackage[font={small}]{caption}
 % Pandoc toggle for numbering sections (defaults to be off)
+
 % Pandoc header
@@
 \usepackage[font={small}]{caption}
-
-\usepackage[nomarkers]{endfloat}
+\usepackage{amsmath}
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{longtable}
 
@@
 
-  \title{A reproducible journal article with RMarkdown: A building science
-template with examples for beginners based on the Elsevier journal
-format}
-    \author[CBE]{Paul Raftery\corref{c1}}
+  \title{A reproducible journal article with RMarkdown: A building science template with examples for beginners based on the Elsevier journal format}
+    \author[CBE]{Paul Raftery\corref{Corresponding Author}}
    \ead{p.raftery@berkeley.edu} 
-   \cortext[c1]{Corresponding Author}
     \author[CBE]{Dana Miller}
@@
 
-      \address[CBE]{Center for the Built Environment, UC Berkeley, 390 Wurster Hall,
-Berkeley, CA, 94720, USA}
+      \address[CBE]{Center for the Built Environment, UC Berkeley, 390 Wurster Hall, Berkeley, CA, 94720, USA}
     \address[Organisation2]{Another organization, and their address}
@@
   \begin{abstract}
-  Enter the text for your abstract here. All data and analysis code in
-  this template is publicly available at
-  \url{https://github.com/CenterForTheBuiltEnvironment/rmd-example} and
-  can be freely adapted and reused. Suggestions or pull requests to
-  improve this template or create additional templates are welcome.
+  Enter the text for your abstract here.
+All data and analysis code in this template is publicly available at \url{https://github.com/CenterForTheBuiltEnvironment/rmd-example} and can be freely adapted and reused.
+Suggestions or pull requests to improve this template or create additional templates are welcome.
   \end{abstract}
@@
 
-\textbf{Graphical Abstract}\\
-\includegraphics{C:/Users/Dana/3_Professional/CBE/Reproducibility/cbe-rmd-example/Paper/SupplementaryMaterial/Images/Latex_logo.png}
+\textbf{Graphical Abstract}
+
+\includegraphics[width=4.44in]{/home/trc/rmd-example/Paper/SupplementaryMaterial/Images/Latex_logo}
 
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-The purpose of this (very much work-in-progress) document is to provide
-a complete R markdown template for an Elsevier journal submission (based
-on the rticles repository Allaire et al.
-(\protect\hyperlink{ref-rticles}{2017})), along with useful examples and
-packages to improve usability for folks who are just starting out with
-this workflow. The eventual intent is to capture minimal examples of the
-common things that authors need to do when writing papers in R markdown;
-provide examples of useful packages, workflows, and tools; and provide
-solutions to common issues that folks encounter.
+The purpose of this (very much work-in-progress) document is to provide a complete R markdown template for an Elsevier journal submission (based on the \href{https://github.com/rstudio/rticles}{rticles} repository Allaire et al.
+(\protect\hyperlink{ref-rticles}{2017})), along with useful examples and packages to improve usability for folks who are just starting out with this workflow.
+The eventual intent is to capture minimal examples of the common things that authors need to do when writing papers in R markdown; provide examples of useful packages, workflows, and tools; and provide solutions to common issues that folks encounter.
 
@@
 
-\section{\texorpdfstring{Methods (this is a `level 1
-heading')}{Methods (this is a level 1 heading)}}\label{methods-this-is-a-level-1-heading}
+\hypertarget{methods-this-is-a-level-1-heading}{%
+\section{Methods (this is a `level 1 heading')}\label{methods-this-is-a-level-1-heading}}
 
-\subsection{\texorpdfstring{Markdown (this is a `level 2
-heading')}{Markdown (this is a level 2 heading)}}\label{markdown-this-is-a-level-2-heading}
+\hypertarget{markdown-this-is-a-level-2-heading}{%
+\subsection{Markdown (this is a `level 2 heading')}\label{markdown-this-is-a-level-2-heading}}
 
-This RMarkdown document is a special type of interactive document that
-can contain both code chunks (in R, python, or other languages), and
-text written in a format called Markdown. Here are some examples of how
-to format text with Markdown, and a
-\href{https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf}{link}
-to an RMarkdown guide.
+This RMarkdown document is a special type of interactive document that can contain both code chunks (in R, python, or other languages), and text written in a format called Markdown.
+Here are some examples of how to format text with Markdown, and a \href{https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf}{link} to an RMarkdown guide.
 
-\subsubsection{Level 3 heading}\label{level-3-heading}
+\hypertarget{level-3-heading}{%
+\subsubsection{Level 3 heading}\label{level-3-heading}}
 
-Here's how to \textbf{bold} or \emph{italicize} a piece of text. This is
-how you do a bullet point list:
+Here's how to \textbf{bold} or \emph{italicize} a piece of text. This is how you do a bullet point list:
 
@@
 
-\subsection{Tables}\label{tables}
+\hypertarget{tables}{%
+\subsection{Tables}\label{tables}}
 
-Here is an example of a table created from the .csv file in the
-\texttt{SupplementaryMaterial/Data}folder using the \texttt{gt} package
-from RStudio. You can read more about \texttt{gt}
-\href{https://github.com/rstudio/gt}{here}, and there's helpful examples
-to further customize tables (eg with color, foototes, significant
-figures, re-labeling columns, and more)
-\href{https://github.com/allisonhorst/gt-awesome-tables}{here}.
+Here is an example of a table created from the .csv file in the \texttt{SupplementaryMaterial/Data}folder using the \texttt{gt} package from RStudio.
+You can read more about \texttt{gt} \href{https://github.com/rstudio/gt}{here}, and there's helpful examples to further customize tables (eg with color, footnotes, significant figures, re-labeling columns, and more) \href{https://github.com/allisonhorst/gt-awesome-tables}{here}.
 
 \captionsetup[table]{labelformat=empty,skip=1pt}
-
 \begin{longtable}{lrrc}
@@
 \midrule
-Iris-virginica & 6.4 & 5.5 & High \\ 
-Iris-virginica & 6.9 & 5.4 & High \\ 
-Iris-setosa & 4.4 & 1.3 & VeryLow \\ 
-Iris-virginica & 7.7 & 6.7 & VeryHigh \\ 
-Iris-versicolor & 5.0 & 3.3 & Medium \\ 
+Iris-setosa & 5.3 & 1.5 & Low \\ 
+Iris-versicolor & 5.6 & 3.6 & Medium \\ 
+Iris-versicolor & 6.1 & 4.7 & High \\ 
+Iris-virginica & 6.7 & 5.2 & High \\ 
+Iris-virginica & 5.6 & 4.9 & High \\ 
 \bottomrule
@@
 
-\subsection{Cross-referencing}\label{cross-referencing}
+\hypertarget{spellcheck}{%
+\subsection{Spellcheck}\label{spellcheck}}
 
-This is how you refer to a figure in your text: Figure
-\ref{fig:correlation}. Simply reference the title of the code chunk, and
-ensure that the code chunk includes a figure caption.
+If you are using RStudio, press F7 or go to \texttt{Edit} --\textgreater{} \texttt{Check\ spelling} in the menu.
 
-\begin{figure}
-\centering
-\includegraphics{C:/Users/Dana/3_Professional/CBE/Reproducibility/cbe-rmd-example/Paper/SupplementaryMaterial/Images/Correlation.png}
-\caption{\label{fig:correlation}Correlation. Source: XKCD, xkcd.com/552}
+\hypertarget{cross-referencing}{%
+\subsection{Cross-referencing}\label{cross-referencing}}
+
+This is how you refer to a figure in your text: Figure \ref{fig:correlation}. Simply reference the title of the code chunk, and ensure that the code chunk includes a figure caption.
+
+\begin{figure}[h]
+\includegraphics[width=0.92in]{/home/trc/rmd-example/Paper/SupplementaryMaterial/Images/correlation} \caption{Correlation. Source: XKCD, xkcd.com/552}\label{fig:correlation}
 \end{figure}
 
-\subsection{Calculations in text}\label{calculations-in-text}
+\hypertarget{calculations-in-text}{%
+\subsection{Calculations in text}\label{calculations-in-text}}
 
-The holy grail of markdown - doing all of your calculations in the same
-file so you never need to worry about updating them after
-someone\footnote{Often I'm the someone, sorry CBE grad students. Also,
-  look, it's an example of a footnote!} asks you to make
-changes\ldots{}. again! It's as easy as pi: 3.14. Incidentally,you can
-selectively override the `global' options set at the beginning, to say
-for example, show more decimals: 3.1416.
+The holy grail of markdown - doing all of your calculations in the same file so you never need to worry about updating them after someone\footnote{Often I'm the someone, sorry CBE grad students.
+Also, look, it's an example of a footnote!} asks you to make changes\ldots{}.
+again!
+It's as easy as pi: 3.14.
+Incidentally,you can selectively override the `global' options set at the beginning, to say for example, show more decimals: 3.1416.
 
-This is an example of outputing the result of a calculation that you
-perform within a code chunk in the document somewhere prior to the
-location where you first refer to it: 36.
+This is an example of outputing the result of a calculation that you perform within a code chunk in the document somewhere prior to the location where you first refer to it: 36.
 
-\subsection{Citations}\label{citations}
+\hypertarget{citations}{%
+\subsection{Citations}\label{citations}}
 
-Citing other literature is remarkably easy, just like this Coakley,
-Raftery, \& Keane
-(\protect\hyperlink{ref-coakleyReviewMethodsMatch2014}{2014}). This
-citation key references the tag associated with an entry in
-Bibliography.bib (a BibTex file). I've found it easiest to use Zotero to
-manage my library of references and to generate the BibTex file, though
-any software that creates a valid BibTex file should work fine. Zotero
-allows you to create a `Collection' (or folder) that gathers together
-all of the references used for a particular document. When combined with
-with the Better BibTex plugin, that collection can be exported to a
-BibTex file that is continually updated as you add or edit references in
-that Collection. Better BibTex also puts the citation key - the text
-after the `@' sybmbol in the .Rmd file - on the upper right of each
-entry, which is convenient for adding citations.
+\hypertarget{citing-literature}{%
+\subsubsection{Citing literature}\label{citing-literature}}
 
-There's not much else involved in citing, as the references list gets
-built and formatted automatically based on the selected style. The only
-other issue I've had to look around to solve was figuring out how to
-combine multiple citations, which is easy when you know how. (Coakley et
-al., \protect\hyperlink{ref-coakleyReviewMethodsMatch2014}{2014}; Zhai
-et al., \protect\hyperlink{ref-zhaiHumanComfortPerceived2015a}{2015})
+Citing other literature is remarkably easy, just like this Coakley et al.
+(\protect\hyperlink{ref-coakleyReviewMethodsMatch2014}{2014}).
+This citation key references the tag associated with an entry in Bibliography.bib (a BibTex file).
+I've found it easiest to use Zotero to manage my library of references and to generate the BibTex file, though any software that creates a valid BibTex file should work fine.
+Zotero allows you to create a `Collection' (or folder) that gathers together all of the references used for a particular document.
+When combined with with the Better BibTex plugin, that collection can be exported to a BibTex file that is continually updated as you add or edit references in that Collection.
+Better BibTex also puts the citation key - the text after the `@' symbol in the .Rmd file - on the upper right of each entry, which is convenient for adding citations.
 
-The instructions for installing BetterBibtex are
-\href{https://retorque.re/zotero-better-bibtex/}{here}.
+There's not much else involved in citing, as the references list gets built and formatted automatically based on the selected style.
+The only other issue I've had to look around to solve was figuring out how to combine multiple citations, which is easy when you know how.
+(Coakley et al., \protect\hyperlink{ref-coakleyReviewMethodsMatch2014}{2014}; Zhai et al., \protect\hyperlink{ref-zhaiHumanComfortPerceived2015a}{2015})
 
-\subsubsection{Citing software}\label{citing-software}
+\textbf{Step-by-step instructions for creating a new Zotero collection and adding a citation}
+1. Install BetterBibtex with the instructions \href{https://retorque.re/zotero-better-bibtex/}{here}.
 
-It is also good practice to include citations not such for literature
-like journal articles and reports, but also the software used in the
-analysis. Many software packages have accompanying publications that can
-be cited, and even if they don't it is important to attribute which
-version and which software you used. Citing software helps to provide
-credit for the creators and maintainors of the software and demonstrate
-that people (like you!) are using the software, which is especially
-important for open-source tools supported through public research
-funding.
-
-You can use the \texttt{greatful} package to quickly generate a list of
-the packages for software used in an R project using the
-\texttt{cite\_packages} function from \texttt{greatful}. There is an
-additional function called \texttt{get\_citations} that generates the
-Bibtex-formatted citations for each package. It's not elegant, but you
-can use the first to list packages used, then generate the
-\texttt{pkg-refs.bib} file (saved to
-\texttt{/Paper/SuplementaryMaterial} by default in this template), copy
-the content into your overall \texttt{Bibliography.bib} file for the
-whole document, and then cite the packages in your document as you would
-another source, eg by using \texttt{@base} for an in-text citation for
-core R software before other packages are installed (also known as `base
-R').
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\item
+  First, create a new Zotero collection for this paper in your own Zotero library.
+You can add existing references by dragging and dropping them into the folder for this collection, or add new references from the web with this collection highlighted.
+\item
+  In the far left Zotero pane, right-click on the folder for this collection, select ``Export Collection'', and selection ``Better BibTeX'' and the format, and tick the ``Keep updated'' box (this will automatically update the stored .bib file when you add new items to this collection).
+Save this .bib file in a location together with the document you are working on (in this repository, it's under \texttt{/Paper})
 
@@
   \item
-  base R - R Core Team (\protect\hyperlink{ref-base}{2018})
+    Note - the dynamic updating \emph{only} works if you export the whole collection to one .bib file, so clicking on the items in the collection and exporting them individually or together will not work
+  \end{itemize}
 \item
-  ggpmisc - Aphalo (\protect\hyperlink{ref-ggpmisc}{2016})
+  To add a citation in the text, first make sure that the item is saved in the Zotero collection that is linked to the automatically-updating .bib file for your document.
+Then click on that item in Zotero and copy the citation key in the right-hand pane (eg \emph{schiavonThermalComfortPerceived2017})
 \item
-  grateful - Rodriguez-Sanchez (\protect\hyperlink{ref-grateful}{2018})
+  In your document, paste the citation key preceded by an @ sign, like this: \texttt{@schiavonThermalComfortPerceived2017} in the text where you want the citation to appear.
+If you want it wrapped in parentheses, do this \texttt{{[}@schiavonThermalComfortPerceived2017{]}}
+\end{enumerate}
+
+\hypertarget{citing-software}{%
+\subsubsection{Citing software}\label{citing-software}}
+
+It is also good practice to include citations for the software used in your analysis.
+After all, the software is also one of the tools you used to carry out your research.
+It is important to attribute which software and which versions you used you used so other people can better understand your methods.
+There have been cases where running the same analysis code with different versions of the same software produces different results!
+In addition, citing software helps to provide credit for the creators and maintainers of the software and demonstrate that people are using it, which is especially important for open-source tools supported through public research funding.
+Just like software often includes a \texttt{LICENSE} file, a good practice is for citation instructions to be included in a \texttt{CITATION} file associated with the project.
+Many software packages have accompanying publications that can be cited, and if they don't there is usually a project website or code repository.
+
+\textbf{Instructions for adding software citations with \texttt{grateful} and Zotero}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Example of generating BibTex file for the packages listed below}
+\CommentTok{# Note the output location is specified }
+\CommentTok{# Note tidyverse is not included in the list below because it's a large package to install on Binder, if you are running this on your own computer I recommend including it if you want to cite it}
+\NormalTok{pkgs <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"here"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"ggpmisc"}\NormalTok{, }\StringTok{"gt"}\NormalTok{, }\StringTok{"grateful"}\NormalTok{, }\StringTok{"base"}\NormalTok{, }\StringTok{"rticles"}\NormalTok{, }\StringTok{"bookdown"}\NormalTok{, }
+          \StringTok{"rmarkdown"}\NormalTok{)}
+\NormalTok{cites <-}\StringTok{ }\NormalTok{grateful}\OperatorTok{::}\KeywordTok{get_citations}\NormalTok{(pkgs, }\DataTypeTok{filename =} \StringTok{"pkg-refs.bib"}\NormalTok{,}
+                    \DataTypeTok{out.dir =}  \KeywordTok{here}\NormalTok{(}\StringTok{"/Paper/SupplementaryMaterial/Software_citations"}\NormalTok{) )}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{4}
 \item
-  gt - Iannone, Cheng, \& Schloerke (\protect\hyperlink{ref-gt}{2019})
+  To add software citations, you can use the \texttt{greatful} package's \texttt{get\_citations} function to generate the BibTeX-formatted citation for each package.
+Note that all of the packages have to be loaded into your working session, and listed before calling \texttt{get\_citations} for this to work.
 \item
-  here - Müller (\protect\hyperlink{ref-here}{2017})
+  Next, open the \texttt{pkg-refs.bib} file generated by \texttt{greatful} in a text editor like Notepad, highlight and copy the text, and then in Zotero with the collection for this project highlighted, click \texttt{File} --\textgreater{} \texttt{Import\ from\ clipboard}.
+Now all the software references have been added to the .bib file that will continue to be updated for this project!
 \item
-  knitr - Xie (\protect\hyperlink{ref-knitr}{2018})
-\item
-  tidyverse - Wickham (\protect\hyperlink{ref-tidyverse}{2017})
-\item
-  rticles - Allaire et al. (\protect\hyperlink{ref-rticles}{2017})
-\item
-  bookdown - Xie (\protect\hyperlink{ref-bookdown}{2016})
-\item
-  rmarkdown - Allaire et al. (\protect\hyperlink{ref-rmarkdown}{2018})
-\end{itemize}
+  Cite the software somewhere in your methods section. Here's an example:
+\end{enumerate}
 
-\subsection{Equations and math}\label{equations-and-math}
+\begin{quote}
+This paper used the free and open source R statistical computing language (R Core Team, \protect\hyperlink{ref-base}{2018}) with \texttt{tidyverse} (Wickham, \protect\hyperlink{ref-tidyverse}{2017}) software for all analysis, with the additional software packages \texttt{ggpmisc} (Aphalo, \protect\hyperlink{ref-ggpmisc}{2016}) for graphics, \texttt{gt} (Iannone et al., \protect\hyperlink{ref-gt}{2019})for tables, \texttt{here} (Müller, \protect\hyperlink{ref-here}{2017}) for file path management, \texttt{grateful} (Rodriguez-Sanchez, \protect\hyperlink{ref-grateful}{2018}) for software citation, \texttt{rmarkdown} (Allaire et al., \protect\hyperlink{ref-rmarkdown}{2018}) for interactive notebooks, and \texttt{knitr} (Xie, \protect\hyperlink{ref-knitr}{2018}), \texttt{rticles} (Allaire et al., \protect\hyperlink{ref-rticles}{2017}), and \texttt{bookdown} (Xie, \protect\hyperlink{ref-bookdown}{2016}) to create a journal-formatted PDF and other markdown and html files.
+\end{quote}
 
-Here's a basic example inline \(example_{subscript} = \frac{D}{R}\), or
-you display it on a whole line if needed. Google latex math cheat sheets
-for more information.
+\hypertarget{equations-and-math}{%
+\subsection{Equations and math}\label{equations-and-math}}
+
+Here's a basic example inline \(example_{subscript} = \frac{D}{R}\), or you display it on a whole line if needed. Google latex math cheat sheets for more information.
 
@@
 
-\subsection{Other packages}\label{other-packages}
+\hypertarget{other-packages}{%
+\subsection{Other packages}\label{other-packages}}
 
-There are lots of packages that are useful for markdown docs and
-customizing plots. We encourage you to search for these whenever you
-encounter a new thing you need to do and to propose an addition to this
-repository accordingly. Some examples to start: \texttt{ggExtra},
-\texttt{gridExtra}, \texttt{RColorBrewer}, \texttt{ggrepel}\ldots{}
+There are lots of packages that are useful for markdown docs and customizing plots.
+We encourage you to search for these whenever you encounter a new thing you need to do and to propose an addition to this repository accordingly.
+Some examples to start: \texttt{ggExtra}, \texttt{gridExtra}, \texttt{RColorBrewer}, \texttt{ggrepel}\ldots{}
 
-\subsection{Writing style}\label{writing-style}
+\hypertarget{writing-style}{%
+\subsection{Writing style}\label{writing-style}}
 
-This is a little off topic for an Rmd example but a convenient place to
-remind our grad students about writing style. In almost all cases,
-active voice is better than passive voice. Several psychological studies
-show that the active voice is more easily understood by readers, and
-that information is more accurately reported by authors when writing in
-active voice. For example, research Klenbort \& Anisfeld
-(\protect\hyperlink{ref-klenbortMarkednessPerspectiveInterpretation1974}{1974})
-has shown that the ``active {[}voice{]} offers a neutral structure for
-conveying information''. Authorship guides for highly regarded journals
-often indicate a preference for the active voice instead of passive:
+This is a little off topic for an Rmd example but a convenient place to remind our grad students about writing style.
+In almost all cases, active voice is better than passive voice.
+Several psychological studies show that the active voice is more easily understood by readers, and that information is more accurately reported by authors when writing in active voice.
+For example, research Klenbort \& Anisfeld (\protect\hyperlink{ref-klenbortMarkednessPerspectiveInterpretation1974}{1974}) has shown that the ``active {[}voice{]} offers a neutral structure for conveying information''.
+Authorship guides for highly regarded journals often indicate a preference for the active voice instead of passive:
 
@@
 \item
-  Nature: ``Nature journals like authors to write in the active voice
-  (`we performed the experiment\ldots{}') as experience has shown that
-  readers find concepts and results to be conveyed more clearly if
-  written directly.''(``Nature - How to write a paper,'' n.d.)
+  Nature: ``Nature journals like authors to write in the active voice (`we performed the experiment\ldots{}') as experience has shown that readers find concepts and results to be conveyed more clearly if written directly.''(\emph{Nature - How to Write a Paper}, n.d.)
 \item
-  Science: ``Use active voice when suitable, particularly when necessary
-  for correct syntax (e.g., `To address this possibility, we constructed
-  a lZap library \ldots{},' not `To address this possibility, a lZap
-  library was constructed\ldots{}').'' Ruben, 2012, \& Am
+  Science: ``Use active voice when suitable, particularly when necessary for correct syntax (e.g., `To address this possibility, we constructed a lZap library \ldots{},' not `To address this possibility, a lZap library was constructed\ldots{}').'' Ruben et al.
 (\protect\hyperlink{ref-rubenHowWriteScientist2012}{2012})
@@
 
-And, on top of all that, you also end up with less text if you write in
-active voice, saving space for useful information and making it easier
-for your readers to understand.
+And, on top of all that, you also end up with less text if you write in active voice, saving space for useful information and making it easier for your readers to understand.
 
-\subsubsection{A fun way to spot passive
-voice:}\label{a-fun-way-to-spot-passive-voice}
+\hypertarget{a-fun-way-to-spot-passive-voice}{%
+\subsubsection{A fun way to spot passive voice:}\label{a-fun-way-to-spot-passive-voice}}
 
-If you can add the words `by zombies' (``A scary-easy way to help you
-find passive voice!''
-\protect\hyperlink{ref-ScaryeasyWayHelp2014}{2014}) to the end of the
-sentence and the sentence still makes logical sense, then the sentence
-is in passive voice. You can also switch on the grammar settings in
-Microsoft Word's spelling and grammar checker and it will show up that
-way.
+If you can add the words `by zombies' (``A Scary-Easy Way to Help You Find Passive Voice!'' \protect\hyperlink{ref-ScaryeasyWayHelp2014}{2014}) to the end of the sentence and the sentence still makes logical sense, then the sentence is in passive voice.
+You can also switch on the grammar settings in Microsoft Word's spelling and grammar checker and it will show up that way.
 
@@
 
-``These measurements are not quantitatively reported in the paper''
-(\ldots{} by zombies)
+``These measurements are not quantitatively reported in the paper'' (\ldots{} by zombies)
 
@@
 
-Or even better, it's really the authors doing the reporting as the paper
-is an inanimate object\ldots{}
+Or even better, it's really the authors doing the reporting as the paper is an inanimate object\ldots{}
 
@@
 
-``Six different table and partition configurations were tested''
-(\ldots{} by zombies)
+``Six different table and partition configurations were tested'' (\ldots{} by zombies)
 
@@
 
-\section{Results}\label{results}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-\subsection{Scatter plot example}\label{scatter-plot-example}
+\hypertarget{scatter-plot-example}{%
+\subsection{Scatter plot example}\label{scatter-plot-example}}
 
@@
 
-\subsection{Violin plot example}\label{violin-plot-example}
+\hypertarget{violin-plot-example}{%
+\subsection{Violin plot example}\label{violin-plot-example}}
 
@@
 
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-\section{Limitations}\label{limitations}
+\hypertarget{limitations}{%
+\section{Limitations}\label{limitations}}
 
-It is important to include a discussion of limitations in any paper.
-Some limitations of this template include:
+It is important to include a discussion of limitations in any paper. Some limitations of this template include:
 
@@
 \item
-  Whether or not figures are included inline with text appears to depend
-  of the version of LaTeX the document is compiled with. Compiling on a
-  computer where TinyTeX is installed (eg in the Binder associated with
-  the reporsitory for this project on GitHub) will render the figures
-  inline, whereas compiling the same document on a Windows computer with
-  MikTeX installed does not
+  Whether or not figures are included inline with text appears to depend of the version of LaTeX the document is compiled with.
+Compiling on a computer where TinyTeX is installed (eg in the Binder associated with the repository for this project on GitHub) will render the figures inline, whereas compiling the same document on a Windows computer with MikTeX installed does not
 \item
@@
 
-\section{Conclusion}\label{conclusion}
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
-\section{Acknowledgements}\label{acknowledgements}
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
 
-Don't forget to acknowledge the funder(s) with associated grant numbers
-if required. The same goes for folks who significantly assisted you with
-this paper but that are not authors. Eg ``Agency (grant number 12345)
-supported this work, with cost share provided by the Center for the
-Built Environment. We thank Person1 and Person2 for assistance with data
-collection''
+Don't forget to acknowledge the funder(s) with associated grant numbers if required.
+The same goes for folks who significantly assisted you with this paper but that are not authors.
+Eg ``Agency (grant number 12345) supported this work, with cost share provided by the Center for the Built Environment.
+We thank Person1 and Person2 for assistance with data collection''
 
-\section{Declaration of interest}\label{declaration-of-interest}
+\hypertarget{declaration-of-interest}{%
+\section{Declaration of interest}\label{declaration-of-interest}}
 
-Descibe any relevant interests of the authors, particularly if there is
-a link to the research that is relatively uncommon and could be
-perceived as a conflict of interest. Otherwise : All authors declare no
-conflict of interest.
+Describe any relevant interests of the authors, particularly if there is a link to the research that is relatively uncommon and could be perceived as a conflict of interest.
+Otherwise : All authors declare no conflict of interest.
 
-\section*{References}\label{references}
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}
@@
 \hypertarget{refs}{}
-\hypertarget{ref-ScaryeasyWayHelp2014}{}
-A scary-easy way to help you find passive voice! (2014). \emph{A
-scary-easy way to help you find passive voice! \textbar{} Grammarly
-Blog}.
+\leavevmode\hypertarget{ref-rticles}{}%
+Allaire, J., R Foundation, Wickham, H., Journal of Statistical Software, Xie, Y., Vaidyanathan, R., Association for Computing Machinery, Boettiger, C., Elsevier, Broman, K., Mueller, K., Quast, B., Pruim, R., Marwick, B., Wickham, C., Keyes, O., \& Yu, M.
+(2017).
+\emph{Rticles: Article formats for r markdown}.
+\url{https://CRAN.R-project.org/package=rticles}
+
+\leavevmode\hypertarget{ref-rmarkdown}{}%
+Allaire, J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins, A., Wickham, H., Cheng, J., \& Chang, W.
+(2018).
+\emph{Rmarkdown: Dynamic documents for r}.
+\url{https://CRAN.R-project.org/package=rmarkdown}
+
+\leavevmode\hypertarget{ref-ggpmisc}{}%
+Aphalo, P. J. (2016). \emph{Learn r ...as you learnt your mother tongue}. Leanpub. \url{https://leanpub.com/learnr}
+
+\leavevmode\hypertarget{ref-ScaryeasyWayHelp2014}{}%
+A scary-easy way to help you find passive voice!
+(2014).
+In \emph{A scary-easy way to help you find passive voice!
+\textbar{} Grammarly Blog}.
 https://www.grammarly.com/blog/a-scary-easy-way-to-help-you-find-passive-voice/.
 
-\hypertarget{ref-rticles}{}
-Allaire, J., R Foundation, Wickham, H., Journal of Statistical Software,
-Xie, Y., Vaidyanathan, R., \ldots{} Yu, M. (2017). \emph{Rticles:
-Article formats for r markdown}. Retrieved from
-\url{https://CRAN.R-project.org/package=rticles}
+\leavevmode\hypertarget{ref-coakleyReviewMethodsMatch2014}{}%
+Coakley, D., Raftery, P., \& Keane, M.
+(2014).
+A review of methods to match building energy simulation models to measured data.
+\emph{Renewable and Sustainable Energy Reviews}, \emph{37}(Supplement C), 123--141.
+\url{https://doi.org/10.1016/j.rser.2014.05.007}
 
-\hypertarget{ref-rmarkdown}{}
-Allaire, J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
-A., \ldots{} Chang, W. (2018). \emph{Rmarkdown: Dynamic documents for
-r}. Retrieved from \url{https://CRAN.R-project.org/package=rmarkdown}
+\leavevmode\hypertarget{ref-gt}{}%
+Iannone, R., Cheng, J., \& Schloerke, B. (2019). \emph{Gt: Easily create presentation-ready display tables}. \url{https://github.com/rstudio/gt}
 
-\hypertarget{ref-ggpmisc}{}
-Aphalo, P. J. (2016). \emph{Learn r ...as you learnt your mother
-tongue}. Leanpub. Retrieved from \url{https://leanpub.com/learnr}
-
-\hypertarget{ref-coakleyReviewMethodsMatch2014}{}
-Coakley, D., Raftery, P., \& Keane, M. (2014). A review of methods to
-match building energy simulation models to measured data.
-\emph{Renewable and Sustainable Energy Reviews}, \emph{37}(Supplement
-C), 123--141. \url{https://doi.org/10.1016/j.rser.2014.05.007}
-
-\hypertarget{ref-gt}{}
-Iannone, R., Cheng, J., \& Schloerke, B. (2019). \emph{Gt: Easily create
-presentation-ready display tables}. Retrieved from
-\url{https://github.com/rstudio/gt}
-
-\hypertarget{ref-klenbortMarkednessPerspectiveInterpretation1974}{}
-Klenbort, I., \& Anisfeld, M. (1974). Markedness and perspective in the
-interpretation of the active and passive voice. \emph{Quarterly Journal
-of Experimental Psychology}, \emph{26}(2), 189--195.
+\leavevmode\hypertarget{ref-klenbortMarkednessPerspectiveInterpretation1974}{}%
+Klenbort, I., \& Anisfeld, M.
+(1974).
+Markedness and perspective in the interpretation of the active and passive voice.
+\emph{Quarterly Journal of Experimental Psychology}, \emph{26}(2), 189--195.
 \url{https://doi.org/10.1080/14640747408400404}
 
-\hypertarget{ref-here}{}
-Müller, K. (2017). \emph{Here: A simpler way to find your files}.
-Retrieved from \url{https://CRAN.R-project.org/package=here}
+\leavevmode\hypertarget{ref-here}{}%
+Müller, K. (2017). \emph{Here: A simpler way to find your files}. \url{https://CRAN.R-project.org/package=here}
 
-\hypertarget{ref-NatureHowWrite}{}
-Nature - How to write a paper. (n.d.).
-https://www.nature.com/authors/author\_resources/how\_write.html.
+\leavevmode\hypertarget{ref-NatureHowWrite}{}%
+\emph{Nature - How to write a paper}. (n.d.). https://www.nature.com/authors/author\_resources/how\_write.html.
 
-\hypertarget{ref-base}{}
-R Core Team. (2018). \emph{R: A language and environment for statistical
-computing}. Vienna, Austria: R Foundation for Statistical Computing.
-Retrieved from \url{https://www.R-project.org/}
+\leavevmode\hypertarget{ref-base}{}%
+R Core Team. (2018). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}
 
-\hypertarget{ref-grateful}{}
-Rodriguez-Sanchez, F. (2018). \emph{Grateful: Facilitate citation of r
-packages}. Retrieved from \url{https://github.com/Pakillo/grateful}
+\leavevmode\hypertarget{ref-grateful}{}%
+Rodriguez-Sanchez, F. (2018). \emph{Grateful: Facilitate citation of r packages}. \url{https://github.com/Pakillo/grateful}
 
-\hypertarget{ref-rubenHowWriteScientist2012}{}
-Ruben, A., 2012, \& Am, 8. (2012). How to Write Like a Scientist.
-\emph{Science \textbar{} AAAS}.
-https://www.sciencemag.org/careers/2012/03/how-write-scientist.
+\leavevmode\hypertarget{ref-rubenHowWriteScientist2012}{}%
+Ruben, A., 2012, \& Am, 8. (2012). How to Write Like a Scientist. In \emph{Science \textbar{} AAAS}. https://www.sciencemag.org/careers/2012/03/how-write-scientist.
 
-\hypertarget{ref-tidyverse}{}
-Wickham, H. (2017). \emph{Tidyverse: Easily install and load the
-'tidyverse'}. Retrieved from
-\url{https://CRAN.R-project.org/package=tidyverse}
+\leavevmode\hypertarget{ref-tidyverse}{}%
+Wickham, H. (2017). \emph{Tidyverse: Easily install and load the 'tidyverse'}. \url{https://CRAN.R-project.org/package=tidyverse}
 
-\hypertarget{ref-bookdown}{}
-Xie, Y. (2016). \emph{Bookdown: Authoring books and technical documents
-with R markdown}. Boca Raton, Florida: Chapman; Hall/CRC. Retrieved from
-\url{https://github.com/rstudio/bookdown}
+\leavevmode\hypertarget{ref-bookdown}{}%
+Xie, Y. (2016). \emph{Bookdown: Authoring books and technical documents with R markdown}. Chapman; Hall/CRC. \url{https://github.com/rstudio/bookdown}
 
-\hypertarget{ref-knitr}{}
-Xie, Y. (2018). \emph{Knitr: A general-purpose package for dynamic
-report generation in r}. Retrieved from \url{https://yihui.name/knitr/}
+\leavevmode\hypertarget{ref-knitr}{}%
+Xie, Y. (2018). \emph{Knitr: A general-purpose package for dynamic report generation in r}. \url{https://yihui.name/knitr/}
 
-\hypertarget{ref-zhaiHumanComfortPerceived2015a}{}
+\leavevmode\hypertarget{ref-zhaiHumanComfortPerceived2015a}{}%
 Zhai, Y., Zhang, Y., Zhang, H., Pasut, W., Arens, E., \& Meng, Q.
-(2015). Human comfort and perceived air quality in warm and humid
-environments with ceiling fans. \emph{Building and Environment},
-\emph{90}, 178--185.
+(2015).
+Human comfort and perceived air quality in warm and humid environments with ceiling fans.
+\emph{Building and Environment}, \emph{90}, 178--185.
 \url{https://doi.org/10.1016/j.buildenv.2015.04.003}

### 2020-06-12 Paper/Manuscript.Rmd
Re-knit w/ fixing name of graphic
New title!

@@
 ---
-title: A reproducible journal article with RMarkdown&#58; A  building science template with examples for beginners based on the Elsevier journal format 
+title: A big change! 
 author:
@@
 ```{r correlation, fig.width = 1, fig.cap = "Correlation. Source: XKCD, xkcd.com/552"}
-include_graphics(here("Paper","SupplementaryMaterial","Images","Correlation.png"), dpi= 500)
+include_graphics(here("Paper","SupplementaryMaterial","Images","correlation.png"), dpi= 500)
 ```

### 2023-08-28 journal-article-pdf/Paper/Manuscript.tex
added a word report example

--- 2020-06-12 Paper/Manuscript.tex
@@
             pdfauthor={},
-            pdftitle={A reproducible journal article with RMarkdown: A building science template with examples for beginners based on the Elsevier journal format},
+            pdftitle={A big change!},
             colorlinks=true,
@@
 
-  \title{A reproducible journal article with RMarkdown: A building science template with examples for beginners based on the Elsevier journal format}
+  \title{A big change!}
     \author[CBE]{Paul Raftery\corref{Corresponding Author}}

### 2024-10-09 pdf (outdated circa 2015)/Paper/SupplementaryMaterial/src/SupplementaryMaterialsScript.R
removing ancient pdf example

@@ -1,2 +1,0 @@
-# Script for performing preprocessing (or other) calculations 
-# that take too long to run conveniently within an R markdown knit process.

### 2024-10-09 pdf (outdated circa 2015)/install.R

@@ -1,15 +1,0 @@
-install.packages("magrittr")
-install.packages("dplyr")
-install.packages("readr")
-install.packages("ggplot2")
-install.packages("rmarkdown")
-install.packages("ggpmisc")
-install.packages("here")
-install.packages("knitr")
-install.packages("bookdown")
-install.packages("rticles")
-install.packages("devtools")
-devtools::install_github("rstudio/gt")
-devtools::install_github("Pakillo/grateful")
-install.packages("tinytex")
-tinytex::install_tinytex()

### 2024-10-09 word/example.qmd
added blank template

---
title: "Title of the example word file"
author:
  - Paul Raftery ^[Center for the Built Environment, University of California Berkeley, USA] ^[Correspondence to p.raftery@berkeley.edu], Author2AtSameOrg ^1^, Author3atDifferentOrg ^[DifferentOrg]
date: "`r Sys.Date()`"
output:
  bookdown::word_document2:
    reference_docx: "./template.docx" # this is the template containing the formatting for the Word file (font of headings, etc.)
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib # this points to a file containing your references for this paper
csl : elsevier-with-titles.csl # you can change citation style by changing this to refer to a different CSL file style
---

```{r knitrsetup, include = FALSE, cache = FALSE}
# knitr setup
# these are the defaults for how knitr reads and processes this file
# you can override them for specific chunks in the qmd file.
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F,
  dev = "jpeg",
  cache = T,
  dpi = 300,
  fig.path = "figs",
  fig.show = "hold",
  fig.pos = "b",
  fig.width = 6.5, fig.height = 8.5, # these are the default figure sizes
  fig.path = "./word"
)

# str(knitr::opts_chunk$get()) # run this in the console for all knitr config options
```

```{r librariesandthemesettings, include = FALSE, cache = FALSE}
# load all required libraries and configure themes

require(pacman)

# load packages using pacman
pacman::p_load(
  tidyverse, lubridate, here, rmarkdown, scales, patchwork, magrittr, janitor, # general, commonly used
  styler, # auto style your code with this addin
  # ggpubr, # ggplot extension focuses on annotations and fitted models
  # ggExtra, # ggplot extension handy for things like marginal density curves on figures
  # sjstats, pwr, # anova results
  # slider, # moving averages
  # rstatix, #pipe friendly stats
  palmerpenguins # quarto functionality
)

options(
  scipen = 999, # turn off scientific notation,
  digits = 2, # set default significant digits to 2, note this only works after a decimal point, use round() or format() to reduce precision of larger numbers
  timeout = 60
)

# set directory
here::i_am("example.qmd")

# ## for example, theme minimal
# # set default theme for ggplot
# theme_set(theme_minimal())
#
# # modify as preferred, e.g.
# theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
#              plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
#              plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
#              plot.background = element_rect(fill = "white", colour = NA),
#              panel.grid.minor = element_blank(),
#              panel.grid.major = element_blank(),
#              axis.text = element_text(size = 10),
#              strip.text = element_text(size = 10, color = "grey20", face = "bold"),
#              strip.background = element_blank())

# or theme light, with some  Cal colors
# figure theme setup
# set base font size for plots
theme_set(theme_light(base_size = 10))
# cleaner display of 'strips' above facets in plots
theme_update(
  strip.background = element_rect(fill = "#F5F5F5", color = "grey"),
  strip.text = element_text(color = "black"),
  strip.text.y = element_text(angle = 90)
)
#  lighten gridlines so just visible
theme_update(panel.grid = element_line(linetype = "dotted"))
# define a reasonably print and color-blind friendly color palette
# these also conveniently happen to be colors from the official UC Berkeley color palette. https://brand.berkeley.edu/identity-and-visual-design/color
cal_palette2 <- c("#3B7EA1", "#C4820E")
# with slight mod to dark yellow for better grayscale performance for 4 color scale
cal_palette4 <- c("#003262", "#b0750c", "#3b7ea1", "#fdb515")
# adding 5th color
cal_palette5 <- c("#003262", "#b0750c", "#3b7ea1", "#fdb515", "#D9661F")
```

```{r functions}
# user defined functions specific to this analysis
```

```{r dataprep, include = FALSE}
# Read in data, misc initial prep/cleaning tasks
# for example, list all the files in a folder that match a specific pattern (e.g. csv files)

files <- list.files(path = "./input_file_folder/", pattern = "*.csv")

# create a short function to read in those files one by one
# modify those files after you read them in,
# and return a dataframe
read_and_append_file <- function(fle) {
  df <- read_csv(str_c("./input_file_folder/", fle))
  # do whatever individual file level modifications you need here
  # for example, extract some string from the filename
  sss <- str_extract(string = fle, pattern = "(?<=df_).*(?=_USA_Clima)")
  df %<>% mutate(sourcefilename = fle, stringofinterest = sss)
  return(df)
}

# apply that function and return an assembled dataframe
df <- files %>% map_dfr(read_and_append_file)

# apply functions to tidy up the dataframe as necessary
df %<>% clean_names()
```

# Abstract {.unnumbered}
The above heading doesn't count towards heading numbers. This is handy for abstracts, nomenclature, etc. If you want to change the style and formatting of any of the headings in the finished document you can do so by modifying the Styles in the word template referenced above, 'template.docx'.

# First level heading
This is how you cite a reference [@rafteryQuantifyingEnergyLosses2018]. It must be in the bibliography file 'references.bib', which is in the bibtex file format and can be automatically output by a citation management software.

## Second level heading
You can and should perform any short calculations directly in the text. For example, there are `r nrow(iris)` rows in the Iris dataset. The mean  species length is `r mean(iris$Sepal.Length)`. Anything more substantial should be calculated (and well commented) in a separate code chunk. Note the number of significant digits is set globally above but can be overwritten as needed, for example that same value to more precision is `r format(mean(iris$Sepal.Length),digits=5)`. You can also reduce precision in large values above the decimal where needed, for example, report `r 10000*mean(iris$Sepal.Length)` as `r round(10000*mean(iris$Sepal.Length),-3)`.

```{r}
# you'll often need to perform calculations that require multiple lines of code and should be commented
# you can do these in a separate code chunk like this one
# any variables you define here can be used in any subsequent point in the doc
# that applies both to other chunks

weird_iris <- iris # create a version of the iris dataset with some random noise added to one of the results columns
weird_iris %<>%
  mutate(Sepal.Length = Sepal.Length * rnorm(nrow(weird_iris), mean = 1, sd = 0.1)) # add a multiplier of random noise of 0.1 standard deviation
```

The mean species length of the modified iris dataset is `r mean(weird_iris$Sepal.Length)`.

### Third level heading
Here's how you cross reference a figure \@ref(fig:yourfignamehere), or a table \@ref(tab:yourtablenamehere) or a section. This is how you include a footnote[^1].

[^1]: Your footnote text goes here. It can go anywhere in the .qmd file but it seems easiest to put it at the end of the paragraph where the footnote is first included.

```{r yourfignamehere, fig.cap="Your fig caption here", fig.width=6.5, fig.height=3}
# note that your fig name should ONLY USE letters OR NUMBERS as there are exceptional hard to debug issues that happen when you use symbols
#
# note that you should define the width and height in inches of your figure to fit and be legible on the template page type
# default is letter size, which means a max width of 6.5 inches and a max length of ~8.5inches depending on caption size
# another option is a single column size, which is a width of ~ 3.5 inches
p1 <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point()

p2 <- ggplot(weird_iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point()

# you can directly output a plot or multiple plots from this code chunk. however,
# often if you have more than one it's best to decide on how you want them organized,
# add tags so you can refer to sub plots in the text, etc. see below
p1 + p2 + plot_layout(ncol = 2) +
  plot_annotation(tag_levels = c("a"), tag_suffix = ")") &
  theme(
    plot.tag.position = c(0, 1),
    plot.tag = element_text(color = "darkred")
  )
```

```{r yourtablenamehere, echo=FALSE}
knitr::kable(mtcars[1:5, 1:5], caption = "Your table caption here")
```

You can include an equation using LaTeX code, 
for example see equation \@ref(eq:exampleequation).

\begin{equation}
    \theta^{*}|L^{*}, \delta^{2},Y,\sigma^{2}_{u} \sim N\left(B^{-1}W'L^{*},\delta^{2}B^{-1} \right)
    (\#eq:exampleequation)
\end{equation}

#### Fourth level heading (but don't use this many)
No need to deal with specifying superscript or subscript as you can just use 
the relevant unicode symbol directly in either the body text or in code chunks, 
e.g. ft², m³, CO₂, ½, °C.

<!-- You can copy paste in many more subscript/superscript symbols from here: 
https://en.wikipedia.org/wiki/Unicode_subscripts_and_superscripts and there 
are even more symbols here: https://en.wikipedia.org/wiki/Unicode_block. Also, this is how you add a comment in your body text. -->

# Aiding the review process
Though the output is a formatted Word file, the review process is still a bit 
clunky. I generally turn on track changes and send to co-authors for review. 
Then you manually bring those changes back into the .qmd file.  
Sometimes it's handy to add a 
[That way you can comment on why something was done a particular way, 
or propose an alternate, etc.,and folks can respond 
in that thread]{.comment-start id="0" author="The Flying Spahetti Monster" date="2300-01-01T10:12:00Z"}comment in the actual Word file output.[]{.comment-end id="0"}

# CRediT authorship contribution statement {.unnumbered}
**Author1Name**: Conceptualization, Data curation, Formal analysis,
Funding acquisition, Methodology, Investigation, Project administration,
Software, Supervision, Writing - original draft, Writing - review &
editing.

# Declaration of competing interest {.unnumbered}
All authors declare XYZ competing interest. The Center for the Built Environment (CBE) at the University of California, Berkeley, with which one authors is affiliated, is advised and funded in part by many approximately 50 partners that represent a diversity of organizations from the building industry, including manufacturers, building owners, facility managers, contractors, architects, engineers, government agencies, and utilities.

# Acknowledgements {.unnumbered}
The Funding Agency (Grant: XYZ) and the Center for the
Built Environment at the University of California, Berkeley supported
this research. Don't forget to thank any other folks who contributed!

# References {#references .unnumbered}
The string used to cite a reference is what's used to link the Bibtex file (references.bib) to a specific reference. You can use the BetterBibtex plugin within Zotero to automatically create these for your project, and to write the associated .bib file.

### 2024-10-18 word/blank.qmd
tidy to only required files

---
title: "TODOTitle"
author:
  - TODOAuthor1 ^[Center for the Built Environment, University of California Berkeley, USA] ^[Correspondence to TODO@berkeley.edu], TODOAuthor2AtSameOrg ^1^, TODOAuthor3atDifferentOrg ^[TODODifferentOrg]
date: "`r Sys.Date()`"
output:
  bookdown::word_document2:
    reference_docx: "./template.docx" 
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib 
csl : elsevier-with-titles.csl 
---

```{r knitrsetup, include = FALSE, cache = FALSE}
# knitr setup
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F,
  dev = "jpeg",
  cache = T,
  dpi = 300,
  fig.path = "figs",
  fig.show = "hold",
  fig.pos = "b",
  fig.width = 6.5, fig.height = 8.5, # these are the default figure sizes
  fig.path = "./figs"
)

# str(knitr::opts_chunk$get()) # run this in the console for all knitr config options
```

```{r librariesandthemesettings, include = FALSE, cache = FALSE}
# load all required libraries and configure themes

require(pacman)

# load packages using pacman
pacman::p_load(
  tidyverse, lubridate, here, rmarkdown, scales, patchwork, magrittr, janitor, # general, commonly used
  measurements, # most unit conversions 
  styler, # auto style your code with this addin
  # ggpubr, # ggplot extension focuses on annotations and fitted models
  # ggExtra, # ggplot extension handy for things like marginal density curves on figures
  # sjstats, pwr, # anova results
  # slider, # moving averages
  # rstatix, #pipe friendly stats
  palmerpenguins # quarto functionality
)

options(
  scipen = 999, # turn off scientific notation,
  digits = 2, # set default significant digits to 2, note this only works after a decimal point, use round() or format() to reduce precision of larger numbers
  timeout = 60
)

# set directory
here::i_am("blank.qmd")

# or theme light, with some  Cal colors
# figure theme setup
# set base font size for plots
theme_set(theme_light(base_size = 10))
# cleaner display of 'strips' above facets in plots
theme_update(
  strip.background = element_rect(fill = "#F5F5F5", color = "grey"),
  strip.text = element_text(color = "black"),
  strip.text.y = element_text(angle = 90)
)
#  lighten gridlines so just visible
theme_update(panel.grid = element_line(linetype = "dotted"))
# define a reasonably print and color-blind friendly color palette
# these also conveniently happen to be colors from the official UC Berkeley color palette. https://brand.berkeley.edu/identity-and-visual-design/color
cal_palette2 <- c("#3B7EA1", "#C4820E")
# with slight mod to dark yellow for better grayscale performance for 4 color scale
cal_palette4 <- c("#003262", "#b0750c", "#3b7ea1", "#fdb515")
# adding 5th color
cal_palette5 <- c("#003262", "#b0750c", "#3b7ea1", "#fdb515", "#D9661F")
```

```{r constants}
# user defined constants specific to this analysis
```

```{r functions}
# user defined functions specific to this analysis
```

```{r dataprep, include = FALSE}
# # Read in data, misc initial prep/cleaning tasks
# # for example, list all the files in a folder that match a specific pattern (e.g. csv files)
# 
# files <- list.files(path = "./input_file_folder/", pattern = "*.csv")
# 
# # create a short function to read in those files one by one
# # modify those files after you read them in,
# # and return a dataframe
# read_and_append_file <- function(fle) {
#   df <- read_csv(str_c("./input_file_folder/", fle))
#   # do whatever individual file level modifications you need here
#   # for example, extract some string from the filename
#   sss <- str_extract(string = fle, pattern = "(?<=df_).*(?=_USA_Clima)")
#   df %<>% mutate(sourcefilename = fle, stringofinterest = sss)
#   return(df)
# }
# 
# # apply that function and return an assembled dataframe
# df <- files %>% map_dfr(read_and_append_file)
# 
# # apply functions to tidy up the dataframe as necessary
# df %<>% clean_names()
```

# Abstract {.unnumbered}

# Introduction

## Literature review

# Methods

# Results

# Discussion

## Limitations

# CRediT authorship contribution statement {.unnumbered}
TODO **Author1Name**: Conceptualization, Data curation, Formal analysis,
Funding acquisition, Methodology, Investigation, Project administration,
Software, Supervision, Writing - original draft, Writing - review &
editing.

# Declaration of competing interest {.unnumbered}
TODO All authors declare XYZ competing interest. The Center for the Built Environment (CBE) at the University of California, Berkeley, with which one authors is affiliated, is advised and funded in part by many approximately 50 partners that represent a diversity of organizations from the building industry, including manufacturers, building owners, facility managers, contractors, architects, engineers, government agencies, and utilities.

# Acknowledgements {.unnumbered}
TODO The Funding Agency (Grant: XYZ) and the Center for the
Built Environment at the University of California, Berkeley supported
this research. Don't forget to thank any other folks who contributed!

# References {#references .unnumbered}

### 2024-10-18 word/example.qmd
tidy to only required files

@@
   tidyverse, lubridate, here, rmarkdown, scales, patchwork, magrittr, janitor, # general, commonly used
+  measurements, # most unit conversions 
   styler, # auto style your code with this addin
@@
 
+```{r constants}
+# user defined constants specific to this analysis
+```
+
 ```{r functions}
@@
 # for example, list all the files in a folder that match a specific pattern (e.g. csv files)
-
-files <- list.files(path = "./input_file_folder/", pattern = "*.csv")
-
-# create a short function to read in those files one by one
-# modify those files after you read them in,
-# and return a dataframe
-read_and_append_file <- function(fle) {
-  df <- read_csv(str_c("./input_file_folder/", fle))
-  # do whatever individual file level modifications you need here
-  # for example, extract some string from the filename
-  sss <- str_extract(string = fle, pattern = "(?<=df_).*(?=_USA_Clima)")
-  df %<>% mutate(sourcefilename = fle, stringofinterest = sss)
-  return(df)
-}
-
-# apply that function and return an assembled dataframe
-df <- files %>% map_dfr(read_and_append_file)
-
-# apply functions to tidy up the dataframe as necessary
-df %<>% clean_names()
+# 
+# files <- list.files(path = "./input_file_folder/", pattern = "*.csv")
+# 
+# # create a short function to read in those files one by one
+# # modify those files after you read them in,
+# # and return a dataframe
+# read_and_append_file <- function(fle) {
+#   df <- read_csv(str_c("./input_file_folder/", fle))
+#   # do whatever individual file level modifications you need here
+#   # for example, extract some string from the filename
+#   sss <- str_extract(string = fle, pattern = "(?<=df_).*(?=_USA_Clima)")
+#   df %<>% mutate(sourcefilename = fle, stringofinterest = sss)
+#   return(df)
+# }
+# 
+# # apply that function and return an assembled dataframe
+# df <- files %>% map_dfr(read_and_append_file)
+# 
+# # apply functions to tidy up the dataframe as necessary
+# df %<>% clean_names()
 ```
